---
title: "DEI Direction-of-Change Study: Technical Appendix"
author: "Casey McMahon"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)

library(tidyverse)
library(psych)
library(broom)
library(emmeans)
library(kableExtra)
library(corrr)
library(ggeffects)

```

```{r helpers, include=FALSE}
format_num <- function(x, digits = 2) {
  dplyr::if_else(is.na(x), "", sprintf(paste0("%.", digits, "f"), x))
}

clean_kable <- function(df, caption = NULL, align = NULL) {
  kableExtra::kable(
    df,
    format = "html",
    caption = caption,
    align = align,
    escape = FALSE
  ) %>%
    kableExtra::kable_styling(
      full_width = FALSE,
      position = "center",
      bootstrap_options = c("striped", "hover", "condensed")
    )
}

format_p_text <- function(p) {
  dplyr::case_when(
    is.na(p) ~ "",
    p < .001 ~ "< .001",
    TRUE ~ sprintf("%.3f", p)
  )
}

clean_lm_table <- function(model, caption = NULL) {
  broom::tidy(model) %>%
    dplyr::mutate(
      term = dplyr::recode(
        term,
        "(Intercept)" = "Intercept",
        "condition_labelPro-DEI \u2192 Anti-DEI" = "Condition: Pro-DEI \u2192 Anti-DEI",
        "div_support" = "DEI support",
        "over_poli_r" = "Political orientation",
        "condition_labelPro-DEI \u2192 Anti-DEI:div_support" = "Condition: Pro-DEI \u2192 Anti-DEI \u00d7 DEI support",
        "condition_labelPro-DEI \u2192 Anti-DEI:over_poli_r" = "Condition: Pro-DEI \u2192 Anti-DEI \u00d7 political orientation",
        .default = term
      ),
      b = sprintf("%.2f", estimate),
      SE = sprintf("%.2f", std.error),
      t = sprintf("%.2f", statistic),
      p = format_p_text(p.value)
    ) %>%
    dplyr::select(
      Predictor = term,
      b,
      SE,
      t,
      p
    ) %>%
    kableExtra::kable(
      format = "html",
      caption = caption,
      align = c("l", "c", "c", "c", "c"),
      escape = TRUE
    ) %>%
    kableExtra::kable_styling(
      full_width = FALSE,
      position = "center",
      bootstrap_options = c("striped", "hover", "condensed")
    )
}

clean_emmeans_table <- function(emm_obj, caption) {
  as.data.frame(emm_obj) %>%
    dplyr::mutate(
      dplyr::across(where(is.numeric), ~ round(.x, 2))
    ) %>%
    clean_kable(caption = caption)
}
```

# Overview

This appendix reports analyses for a survey experiment testing how people evaluate organizations that change direction on DEI. Participants read about one organization that either moved from a pro-DEI stance to an anti-DEI stance or from an anti-DEI stance to a pro-DEI stance. The primary outcomes were perceived hypocrisy and genuineness. The study also measured political orientation, DEI support, win-win beliefs about diversity, social dominance orientation, and interest in learning more about the organization.

The public-facing portfolio page uses this study to show that DEI policy changes are interpreted as credibility signals, and that reactions to those changes are shaped by participants' own political and DEI attitudes.

```{r load-cleaned-data}
data <- read_csv("data/dei-direction-change-cleaned.csv")

# Make the contrast direction stable. The anti-DEI -> pro-DEI condition is the reference group.
data <- data %>%
  dplyr::mutate(
    condition_label = factor(
      condition_label,
      levels = c("Anti-DEI to pro-DEI", "Pro-DEI to anti-DEI")
    ),
    over_poli_s = factor(over_poli_s, levels = c("Liberal", "Moderate", "Conservative"))
  )
```

# Sample and data checks

```{r sample-tables, echo=FALSE, results='asis'}
condition_table <- data %>%
  dplyr::count(condition_label, name = "N") %>%
  dplyr::mutate(Percent = sprintf("%.1f%%", 100 * N / sum(N))) %>%
  dplyr::rename(Condition = condition_label)

clean_kable(condition_table, caption = "Participants by experimental condition")


data = data %>% 
  mutate(race_r =
           case_when(race == 1 ~ 1,  
                     race == 2 ~ 2,
                     race == 3 ~ 3,
                     race == 4 ~ 4,
                     race == 5 ~ 5,
                     race == 6 ~ 6,
                     race == 7 ~ 7,
                     race == 8 ~ 8,
                     race == 12 ~ 7,
                     race == 13 ~ 7,
                     race == 14 ~ 7,
                     race == 15 ~ 7,
                     race == 16 ~ 7,
                     race == 23 ~ 7,
                     race == 28 ~ 7,
                     race == 126 ~ 7,
                     race == 148 ~ 7,
                     race == 1256 ~ 7,
                     race == 13458 ~ 7
                     ))

race_table <- data %>%
  dplyr::count(race_r, name = "N") %>%
  dplyr::filter(!is.na(race_r)) %>%
  dplyr::mutate(
    Percent = sprintf("%.1f%%", 100 * N / sum(N)),
    Race = dplyr::recode(
      as.character(race_r),
      "1" = "White",
      "2" = "Asian",
      "3" = "Black",
      "4" = "Hispanic/ Latino/a",
      "5" = "Native American",
      "6" = "Middle Eastern",
      "7" = "Bi/ multiracial",
      "8" = "Another race",
      .default = as.character(race_r)
    )
  ) %>%
  dplyr::select(Race, N, Percent)

clean_kable(race_table, caption = "Participant race")

ideology_table <- data %>%
  dplyr::count(over_poli_s, name = "N") %>%
  dplyr::filter(!is.na(over_poli_s)) %>%
  dplyr::mutate(Percent = sprintf("%.1f%%", 100 * N / sum(N))) %>%
  dplyr::rename(`Political orientation group` = over_poli_s)

clean_kable(ideology_table, caption = "Participants by political orientation group")

age_table <- data %>%
  dplyr::summarise(
    N = sum(!is.na(age)),
    M = mean(age, na.rm = TRUE),
    SD = sd(age, na.rm = TRUE),
    Min = min(age, na.rm = TRUE),
    Max = max(age, na.rm = TRUE)
  ) %>%
  dplyr::mutate(
    dplyr::across(M:Max, ~ sprintf("%.2f", .x))
  )

clean_kable(age_table, caption = "Participant age")
```

# Measures and scale reliability

This section summarizes the primary scales used in the study. Higher hypocrisy scores indicate stronger perceptions that the organization acted hypocritically. Higher genuineness scores indicate stronger perceptions that the organization was genuinely committed to its current DEI stance. Higher DEI support, win-win beliefs, and SDO scores indicate stronger endorsement of those constructs.

```{r scale-reliability, echo=FALSE, results='asis'}
alpha_table <- tibble::tibble(
  Scale = c(
    "DEI support",
    "Hypocrisy",
    "Genuineness",
    "Social dominance orientation",
    "Win-win beliefs about diversity"
  ),
  Alpha = c(
    psych::alpha(data %>% dplyr::select(div_support_1, div_support_2, div_support_3, div_support_4, div_support_5, div_support_6_r, div_support_7_r), check.keys = TRUE)$total$raw_alpha,
    psych::alpha(data %>% dplyr::select(hypocrisy_1, hypocrisy_2, hypocrisy_3, hypocrisy_4, hypocrisy_5), check.keys = TRUE)$total$raw_alpha,
    psych::alpha(data %>% dplyr::select(gen_1, gen_2_r, gen_3_r, gen_4), check.keys = TRUE)$total$raw_alpha,
    psych::alpha(data %>% dplyr::select(sdo_1, sdo_2, sdo_3r, sdo_4r, sdo_5, sdo_6, sdo_7r, sdo_8r), check.keys = TRUE)$total$raw_alpha,
    psych::alpha(data %>% dplyr::select(wwb_1, wwb_2, wwb_3, wwb_4, wwb_5, wwb_6, wwb_7, wwb_8, wwb_9, wwb_10), check.keys = TRUE)$total$raw_alpha
  )
) %>%
  dplyr::mutate(`Cronbach's alpha` = sprintf("%.2f", Alpha)) %>%
  dplyr::select(Scale, `Cronbach's alpha`)

clean_kable(alpha_table, caption = "Scale reliability")
```

```{r scale-descriptives, echo=FALSE, results='asis'}
scale_labels <- c(
  hypocrisy = "Hypocrisy",
  genuineness = "Genuineness",
  div_support = "DEI support",
  pew_r = "DEI is good/bad at work",
  sdo = "Social dominance orientation",
  wwbs = "Win-win beliefs",
  over_poli_r = "Political orientation"
)

scale_vars <- names(scale_labels)

scale_descriptives <- data %>%
  dplyr::select(dplyr::all_of(scale_vars)) %>%
  tidyr::pivot_longer(dplyr::everything(), names_to = "Scale", values_to = "Value") %>%
  dplyr::group_by(Scale) %>%
  dplyr::summarise(
    N = sum(!is.na(Value)),
    M = mean(Value, na.rm = TRUE),
    SD = sd(Value, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::mutate(
    Scale = dplyr::recode(Scale, !!!scale_labels),
    M = sprintf("%.2f", M),
    SD = sprintf("%.2f", SD)
  )

clean_kable(scale_descriptives, caption = "Scale descriptives")
```

```{r scale-correlations, echo=FALSE, results='asis'}
cor_mat <- data %>%
  dplyr::select(dplyr::all_of(scale_vars)) %>%
  stats::cor(use = "pairwise.complete.obs")

cor_mat[upper.tri(cor_mat, diag = TRUE)] <- NA

cor_table <- as.data.frame(round(cor_mat, 2)) %>%
  tibble::rownames_to_column("Scale") %>%
  dplyr::mutate(Scale = dplyr::recode(Scale, !!!scale_labels))

names(cor_table)[-1] <- dplyr::recode(names(cor_table)[-1], !!!scale_labels)

cor_table <- cor_table %>%
  dplyr::mutate(
    dplyr::across(-Scale, ~ dplyr::if_else(is.na(.x), "", sprintf("%.2f", .x)))
  )

clean_kable(cor_table, caption = "Lower-triangle correlations among study variables")
```

# Descriptive differences by direction of change

These descriptive tables show the average outcome ratings in each condition. They provide the simplest version of the study result: whether participants evaluated the pro-DEI to anti-DEI shift differently from the anti-DEI to pro-DEI shift.

```{r outcome-descriptives, echo=FALSE, results='asis'}
outcome_descriptives <- data %>%
  dplyr::select(condition_label, hypocrisy, genuineness, moreinfo_num) %>%
  tidyr::pivot_longer(
    cols = c(hypocrisy, genuineness, moreinfo_num),
    names_to = "Outcome",
    values_to = "Value"
  ) %>%
  dplyr::mutate(
    Outcome = dplyr::recode(
      Outcome,
      hypocrisy = "Hypocrisy",
      genuineness = "Genuineness",
      moreinfo_num = "Interested in learning more"
    )
  ) %>%
  dplyr::group_by(condition_label, Outcome) %>%
  dplyr::summarise(
    N = sum(!is.na(Value)),
    M = mean(Value, na.rm = TRUE),
    SD = sd(Value, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::mutate(
    M = sprintf("%.2f", M),
    SD = sprintf("%.2f", SD)
  ) %>%
  dplyr::rename(Condition = condition_label)

clean_kable(outcome_descriptives, caption = "Outcome descriptives by condition")
```

```{r outcome-plot, echo=FALSE, fig.width=8, fig.height=5}
outcome_plot_df <- data %>%
  dplyr::select(condition_label, hypocrisy, genuineness) %>%
  tidyr::pivot_longer(
    cols = c(hypocrisy, genuineness),
    names_to = "Outcome",
    values_to = "Value"
  ) %>%
  dplyr::mutate(
    Outcome = dplyr::recode(
      Outcome,
      hypocrisy = "Perceived hypocrisy",
      genuineness = "Perceived genuineness"
    )
  ) %>%
  dplyr::group_by(condition_label, Outcome) %>%
  dplyr::summarise(
    N = sum(!is.na(Value)),
    M = mean(Value, na.rm = TRUE),
    SD = sd(Value, na.rm = TRUE),
    SE = SD / sqrt(N),
    ci_lower = M - qt(.975, df = N - 1) * SE,
    ci_upper = M + qt(.975, df = N - 1) * SE,
    .groups = "drop"
  )

change_direction_outcome_plot <- ggplot(
  outcome_plot_df,
  aes(x = condition_label, y = M)
) +
  geom_point(size = 2.8) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = 0.12, linewidth = 0.6) +
  facet_wrap(~ Outcome, scales = "free_y") +
  labs(
    title = "Perceptions differed by the direction of DEI change",
    x = NULL,
    y = "Mean rating"
  ) +
  coord_cartesian(ylim = c(1,7)) +
  scale_y_continuous(name = "Mean rating", expand = c(0,0), breaks=seq(1,7,1)) + # set ticks at 1, 2, 3, 4 +
  theme_classic(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold"),
    axis.text.x = element_text(angle = 20, hjust = 1)
  )

change_direction_outcome_plot
```

# Main models: hypocrisy

The primary hypothesis for perceived hypocrisy is that people may judge the same act of changing direction differently depending on whether the organization moves away from DEI or toward DEI. These models test whether condition predicts hypocrisy and whether that difference depends on participants' DEI support or political orientation.

```{r hypocrisy-models, echo=FALSE, results='asis'}
hyp_condition_model <- lm(hypocrisy ~ condition_label, data = data)
hyp_dei_support_model <- lm(hypocrisy ~ condition_label * div_support + over_poli_r, data = data)
hyp_political_model <- lm(hypocrisy ~ condition_label * over_poli_r + div_support, data = data)

clean_lm_table(hyp_condition_model, caption = "Hypocrisy model with condition only")
clean_lm_table(hyp_dei_support_model, caption = "Hypocrisy model with DEI support moderation")
clean_lm_table(hyp_political_model, caption = "Hypocrisy model with political orientation moderation")
```

```{r hypocrisy-emtrends, echo=FALSE, results='asis'}
hyp_dei_trends <- emmeans::emtrends(
  hyp_dei_support_model,
  specs = ~ condition_label,
  var = "div_support"
)

clean_emmeans_table(hyp_dei_trends, caption = "Association between DEI support and hypocrisy by condition")

hyp_political_trends <- emmeans::emtrends(
  hyp_political_model,
  specs = ~ condition_label,
  var = "over_poli_r"
)

clean_emmeans_table(hyp_political_trends, caption = "Association between political orientation and hypocrisy by condition")
```

```{r hypocrisy-moderation-plot, echo=FALSE, fig.width=8, fig.height=5}
hyp_dei_predictions <- ggeffects::ggpredict(
  hyp_dei_support_model,
  terms = c("div_support [all]", "condition_label")
) %>%
  as.data.frame()

hypocrisy_dei_moderation_plot <- ggplot(
  hyp_dei_predictions,
  aes(x = x, y = predicted, color = group, fill = group)
) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.15, color = NA) +
  geom_line(linewidth = 0.9) +
  labs(
    title = "DEI support shaped perceived hypocrisy",
    x = "DEI support",
    y = "Predicted hypocrisy",
    color = "Condition",
    fill = "Condition"
  ) +
  coord_cartesian(ylim = c(1,7)) +
  scale_y_continuous(name = "Mean rating", expand = c(0,0), breaks=seq(1,7,1)) + # set ticks at 1, 2, 3, 4 +
  theme_classic(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold"),
    legend.position = "top"
  )

hypocrisy_dei_moderation_plot
```

# Main models: genuineness

Genuineness captures whether participants believed the organization was sincerely committed to its current DEI stance. If the pro-DEI to anti-DEI shift is interpreted as more hypocritical, it should also tend to look less genuine.

```{r genuineness-models, echo=FALSE, results='asis'}
gen_condition_model <- lm(genuineness ~ condition_label, data = data)
gen_dei_support_model <- lm(genuineness ~ condition_label * div_support + over_poli_r, data = data)
gen_political_model <- lm(genuineness ~ condition_label * over_poli_r + div_support, data = data)

clean_lm_table(gen_condition_model, caption = "Genuineness model with condition only")
clean_lm_table(gen_dei_support_model, caption = "Genuineness model with DEI support moderation")
clean_lm_table(gen_political_model, caption = "Genuineness model with political orientation moderation")
```

```{r genuineness-emtrends, echo=FALSE, results='asis'}
gen_dei_trends <- emmeans::emtrends(
  gen_dei_support_model,
  specs = ~ condition_label,
  var = "div_support"
)

clean_emmeans_table(gen_dei_trends, caption = "Association between DEI support and genuineness by condition")

gen_political_trends <- emmeans::emtrends(
  gen_political_model,
  specs = ~ condition_label,
  var = "over_poli_r"
)

clean_emmeans_table(gen_political_trends, caption = "Association between political orientation and genuineness by condition")
```

```{r genuineness-moderation-plot, echo=FALSE, fig.width=8, fig.height=5}
gen_dei_predictions <- ggeffects::ggpredict(
  gen_dei_support_model,
  terms = c("div_support [all]", "condition_label")
) %>%
  as.data.frame()

genuineness_dei_moderation_plot <- ggplot(
  gen_dei_predictions,
  aes(x = x, y = predicted, color = group, fill = group)
) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.15, color = NA) +
  geom_line(linewidth = 0.9) +
  labs(
    title = "DEI support shaped perceived genuineness",
    x = "DEI support",
    y = "Predicted genuineness",
    color = "Condition",
    fill = "Condition"
  ) +
  theme_classic(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold"),
    legend.position = "top"
  )

genuineness_dei_moderation_plot
```

# Interest in learning more

Interest in learning more was exploratory. Because relatively few participants selected yes, these analyses should be interpreted cautiously.

```{r moreinfo-models, echo=FALSE, results='asis'}
moreinfo_condition_model <- glm(moreinfo_num ~ condition_label, data = data, family = binomial)
moreinfo_dei_support_model <- glm(moreinfo_num ~ condition_label * div_support + over_poli_r, data = data, family = binomial)

moreinfo_table <- broom::tidy(moreinfo_dei_support_model) %>%
  dplyr::filter(term != "(Intercept)") %>%
  dplyr::mutate(
    Predictor = dplyr::case_when(
      term == "condition_labelPro-DEI to anti-DEI" ~ "Condition: pro-DEI to anti-DEI",
      term == "div_support" ~ "DEI support",
      term == "over_poli_r" ~ "Political orientation (higher = more conservative)",
      term == "condition_labelPro-DEI to anti-DEI:div_support" ~ "Pro-DEI to anti-DEI × DEI support",
      TRUE ~ term
    ),
    b = sprintf("%.2f", estimate),
    SE = sprintf("%.2f", std.error),
    z = sprintf("%.2f", statistic),
    p = format_p_text(p.value),
    OR = sprintf("%.2f", exp(estimate))
  ) %>%
  dplyr::select(Predictor, b, SE, z, p, OR)

clean_kable(moreinfo_table, caption = "Logistic regression predicting interest in learning more", align = c("l", "c", "c", "c", "c", "c"))
```


```{r export-figures, eval=FALSE, include = FALSE}
# Figures for portfolio
ggsave(
  filename = "../assets/img/dei-direction-change-outcomes.png",
  plot = change_direction_outcome_plot,
  width = 8,
  height = 5,
  dpi = 300
)

ggsave(
  filename = "../assets/img/dei-direction-change-hypocrisy-moderation.png",
  plot = hypocrisy_dei_moderation_plot,
  width = 8,
  height = 5,
  dpi = 300
)
```
