---
title: "DEI Symbolic/Substantive Study: Private Cleaning File"
author: "Casey McMahon"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

# Private cleaning file

This file imports the raw Qualtrics export, applies exclusions and data-quality checks, creates cleaned variables, and saves a de-identified analytic dataset for the technical appendix. This file should remain private because it references raw survey data.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)

library(tidyverse)
library(janitor)
library(readr)
```

## Import raw data

Update `raw_file` to the current location of the raw Qualtrics export.

```{r import-raw-data}
raw_file <- "~/Documents/reseach projects for websites/dei rollbacks/P-I Gap/p-i gap pilot sub v. sym 11.12.25/dei P-I gap policy sym v. sub 11.25_November 13, 2025_10.20.csv"

data_raw <- read.csv(raw_file)
```

## Clean column names

```{r clean-column-names}
data <- data_raw %>%
  janitor::clean_names()
```

## Apply basic exclusions

```{r exclusions}
# Keep real study responses only.
data <- data %>%
  dplyr::filter(distribution_channel == "anonymous")

# Keep participants who consented.
data <- data %>%
  dplyr::filter(consent == "1")
```

## Remove duplicate/bot-like geolocations

```{r geolocation-check}
data <- data %>%
  dplyr::mutate(
    longlat = paste(location_longitude, location_latitude, sep = ", ")
  )

data_removed_geolocation <- data %>%
  dplyr::group_by(longlat) %>%
  dplyr::filter(dplyr::n() > 2) %>%
  dplyr::ungroup()

data <- data %>%
  dplyr::group_by(longlat) %>%
  dplyr::filter(dplyr::n() < 3) %>%
  dplyr::ungroup()
```

## Full-stop attention check

```{r full-stop-check}
data <- data %>%
  dplyr::mutate(
    punctuation = tolower(punctuation),
    punctuation_clean = dplyr::recode(
      punctuation,
      "a period" = "period",
      "a period." = "period",
      "peroid" = "period",
      "period" = "period",
      "period sign" = "period",
      "period." = "period",
      "pereiod" = "period",
      ". period" = "period",
      "the mark used to end a statement is called a period." = "period",
      "the punctuation mark. shown above, is a period." = "period",
      '(".") period' = "period",
      "punctuation mark, period" = "period",
      .default = NA_character_,
      .missing = NA_character_
    )
  )

data_removed_attention <- data %>%
  dplyr::filter(punctuation_clean != "period" | is.na(punctuation_clean))

data <- data %>%
  dplyr::filter(punctuation_clean == "period")
```

## Bot check

```{r bot-check}
data_removed_bots <- data %>%
  dplyr::filter(attbot == "14285733")

data <- data %>%
  dplyr::filter(attbot != "14285733" | is.na(attbot))
```

## Remove private or unnecessary columns

```{r remove-private-columns}
private_cols <- c(
  "start_date",
  "end_date",
  "status",
  "ip_address",
  "progress",
  "duration_in_seconds",
  "recorded_date",
  "response_id",
  "recipient_last_name",
  "recipient_first_name",
  "recipient_email",
  "external_reference",
  "location_latitude",
  "location_longitude",
  "distribution_channel",
  "user_language"
)

data <- data %>%
  dplyr::select(-dplyr::any_of(private_cols))
```

## Add participant ID

```{r participant-id}
data <- data %>%
  tibble::rowid_to_column("ppid") %>%
  dplyr::mutate(ppid = ppid + 100)
```

## Recode political orientation

```{r political-orientation}
data <- data %>%
  dplyr::mutate(
    over_poli = as.numeric(over_poli),
    econ_poli = as.numeric(econ_poli),
    social_poli = as.numeric(social_poli),
    polparty = as.numeric(polparty),
    over_poli_r = 8 - over_poli,
    econ_poli_r = 8 - econ_poli,
    social_poli_r = 8 - social_poli,
    # 0 = conservative/more conservative, 1 = liberal/more liberal
    poli_fact = ifelse(over_poli_r >= 4, 0, 1),
    poli_factor = factor(poli_fact, levels = c(0, 1), labels = c("Conservative", "Liberal")),
    party_group = dplyr::case_when(
      polparty < 4 ~ "Republican-leaning",
      polparty > 4 ~ "Democratic-leaning",
      TRUE ~ NA_character_
    )
  )
```

## Recode value ratings and rankings

```{r value-variables}
data <- data %>%
  dplyr::rename(
    div_likert = value_1,
    inc_likert = value_2,
    bel_likert = value_3,
    equit_likert = value_4,
    merit_likert = value_5,
    just_likert = value_6,
    equal_likert = value_7,
    div_rank = q277_1,
    inc_rank = q277_2,
    bel_rank = q277_3,
    equit_rank = q277_4,
    merit_rank = q277_5,
    just_rank = q277_6,
    equal_rank = q277_7
  ) %>%
  dplyr::mutate(
    dplyr::across(
      c(div_likert, inc_likert, bel_likert, equit_likert, merit_likert, just_likert, equal_likert,
        div_rank, inc_rank, bel_rank, equit_rank, merit_rank, just_rank, equal_rank),
      as.numeric
    ),
    principle = rowMeans(
      dplyr::pick(div_likert, inc_likert, bel_likert, equit_likert, merit_likert, just_likert, equal_likert),
      na.rm = TRUE
    ),
    div_eq = rowMeans(dplyr::pick(div_likert, equit_likert), na.rm = TRUE),
    bel_inc = rowMeans(dplyr::pick(inc_likert, bel_likert), na.rm = TRUE),
    bi_rel_de = div_eq - bel_inc
  )
```

## Create policy substantiveness variables

Each policy was rated on a 0–100 slider. Left-anchor versions are reverse-scored so higher values consistently indicate greater perceived substantiveness.

```{r policy-variables}
# Reverse left-anchor versions so 0 = more symbolic and 100 = more substantive.
for (i in 1:25) {
  left_var <- paste0("x", i, "_polsubl_8")
  left_rev <- paste0("x", i, "_polsubl_8_r")
  right_var <- paste0("x", i, "_polsubr_8")
  policy_var <- sprintf("p%02d", i)

  if (left_var %in% names(data)) {
    data[[left_var]] <- as.numeric(data[[left_var]])
    data[[left_rev]] <- 100 - data[[left_var]]
  }

  if (right_var %in% names(data)) {
    data[[right_var]] <- as.numeric(data[[right_var]])
  }

  data[[policy_var]] <- dplyr::coalesce(data[[right_var]], data[[left_rev]])
}
```

## Create symbolic and substantive composites

The symbolic composite uses the policies that were most clearly symbolic in the original analysis. The substantive composite uses the policies that were most clearly substantive.

```{r symbolic-substantive-composites}
data <- data %>%
  dplyr::mutate(
    symbolic = rowMeans(dplyr::pick(p21, p22, p23, p24, p25), na.rm = TRUE),
    substantive = rowMeans(dplyr::pick(p09, p14, p15, p16, p17), na.rm = TRUE),
    sym_rel_sub = substantive - symbolic
  )
```

## Data cleaning checks

```{r cleaning-checks}
list(
  final_n = nrow(data),
  removed_geolocation = nrow(data_removed_geolocation),
  removed_attention = nrow(data_removed_attention),
  removed_bots = nrow(data_removed_bots)
)
```

## Save analytic dataset

Update the path if needed.

```{r save-clean-data}
output_dir <- "data"
if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)

readr::write_csv(data, file.path(output_dir, "dei-symbolic-substantive-cleaned.csv"))
```
