---
title: "Linked Fate Portfolio Plots"
author: "Casey McMahon"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    keep_md: true
    theme: flatly
---

This file generates figures for the Linked Fate at Work portfolio case study. The graphs are used for the public-facing project page and technical appendices.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(boot)
library(car)
library(CGPfunctions)
library(colorBlindness)
library(corrplot)
library(data.table)
library(descriptr)
library(DescTools)
library(EFA.dimensions)
library(emmeans)
library(furniture)
library(ggeffects)
library(ggpubr)
library(ggsignif)
library(ggstatsplot)
library(GLMMadaptive)
library(gridExtra)
library(janitor)
library(jtools)
library(kableExtra)
library(knitr)
library(lattice)
library(lme4)
library(nlme)
library(patchwork)
library(performance)
library(PMCMRplus)
library(psych)
library(RColorBrewer)
library(readxl)
library(rempsyc)
library(Rmisc)
library(rstatix)
library(see)
library(sjmisc)
library(sjPlot)
library(sjstats)
library(stats)
library(tidyverse) #includes ggplot2, dplyr, tidyr, readr, purrr, tibble, stringr, forcats
library(viridis)
library(WriteXLS)
library(stringr)

# Half violin plots
library(PupillometryR)
source("https://gist.githubusercontent.com/benmarwick/2a1bb0133ff568cbe28d/raw/fb53bd97121f7f9ce947837ef1a4c65a73bffb3f/geom_flat_violin.R")
```

## Study 1: Workplace Relationships
```{r s1data, echo=FALSE, include = FALSE}
# Study 1 workplace relationships data

data_l_lfsb_means <- read_csv("~/Dropbox/Research/current projects/diversity gatekeeping/studies + data/org relationships_powered Jun 24/data + analyses/div gatekeep workplace relations July 2024_long_clean.csv")

```
### Study 1: White Women / Women of Color Comparison

```{r s1g1}

data_l_w = subset(data_l_lfsb_means, scale_name == "lfsb_woc_ww_mean")
data_l_w = subset(data_l_w, gen_name == "women")
data_l_w$group2 = factor(data_l_w$group2, levels = c("Women of color", "White women" )) 

data_l_w <- data_l_lfsb_means %>%
  dplyr::filter(
    scale_name == "lfsb_woc_ww_mean",
    gen_name == "women"
  ) %>%
  dplyr::mutate(
    scale_val = as.numeric(scale_val),
    participant_group = dplyr::case_when(
      race_name == "poc" ~ "Women of color",
      race_name == "white" ~ "White women",
      TRUE ~ NA_character_
    ),
    participant_group = factor(
      participant_group,
      levels = c("Women of color", "White women")
    ),
    x_base = as.numeric(participant_group)
  ) %>%
  dplyr::filter(
    !is.na(participant_group),
    !is.na(scale_val)
  )

ggplot(data_l_w, aes(y = scale_val)) +
  geom_flat_violin(
    aes(
      x = x_base + 0.18,
      group = participant_group,
      fill = participant_group
    ),
    alpha = 0.6,
    color = NA,
    width = 0.55
  ) +
  geom_point(
    aes(x = x_base - 0.08),
    position = position_jitter(width = 0.15, height = 0),
    size = 1,
    alpha = 0.6,
    color = "black",
    show.legend = FALSE
  ) +
  stat_summary(
    aes(
      x = x_base - 0.08,
      group = participant_group
    ),
    fun = mean,
    geom = "point",
    size = 5,
    shape = 18,
    color = "black"
  ) +
  stat_summary(
    aes(
      x = x_base - 0.08,
      group = participant_group
    ),
    fun.data = mean_cl_boot,
    geom = "errorbar",
    width = 0.08,
    linewidth = 1,
    color = "black"
  ) +
  coord_cartesian(ylim = c(1, 7)) +
  scale_y_continuous(
    name = "Linked fate",
    expand = c(0, 0),
    breaks = seq(1, 7, 1)
  ) +
  scale_x_continuous(
    name = NULL,
    breaks = c(1, 2),
    labels = c("Women of color", "White women"),
    limits = c(0.5, 2.6)
  ) +
  scale_fill_manual(
    values = c(
      "Women of color" = "#2FB3EA",
      "White women" = "#89939C"
    )
  ) +
  theme_classic() +
  theme(
    text = element_text(color = "black"),
    axis.text.x = element_text(size = 14, color = "black"),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_text(size = 14, color = "black"),
    axis.title.y = element_text(size = 18),
    legend.position = "none",
    plot.margin = margin(.5, .1, 0, .25, "cm")
  )

```

### Study 1: All Race-Gender Comparisons
#### Figure
```{r s1g2}

pd = position_dodge(.92)

data_l_lfsb_means %>%
  ggplot(aes(x = scale_name, y= scale_val, colour = group2, fill = group2)) +
  geom_bar(stat = "summary", 
           fun = mean,
           position = "dodge") +
  stat_summary(fun.data=mean_se,
               geom='errorbar',
               #fun.args=list(conf.int=.95),
               size=.6,
               aes(width=.3),
               color="black",
               position=pd) +
  coord_cartesian(ylim = c(1,7)) +
  scale_y_continuous(name = "Linked fate", expand = c(0,0), breaks=seq(1,7,1)) + # set ticks at 1, 2, 3, 4 +
  scale_x_discrete(name = "Relationship",
                   limits = c("lfsb_woc_moc_mean",
                        "lfsb_woc_ww_mean",
                        "lfsb_woc_wm_mean",
                        "lfsb_moc_ww_mean",
                        "lfsb_moc_wm_mean",
                        "lfsb_ww_wm_mean"),
                   labels = c("lfsb_woc_moc_mean" = "Women of color/\nMen of color",
                        "lfsb_woc_ww_mean" = "Women of color/\nWhite women",
                        "lfsb_woc_wm_mean" = "Women of color/\nWhite men",
                        "lfsb_moc_ww_mean" = "Men of color/\nWhite women",
                        "lfsb_moc_wm_mean" = "Men of color/\nWhite men",
                        "lfsb_ww_wm_mean" = "White women/\nWhite men")) +
  theme_classic() +
  theme(text = element_text(color = "black"),
        axis.text.x = element_text(color = "black", angle = 45, hjust = 1),
        #axis.title.x = element_blank(),
        axis.text.y = element_text(color = "black"),
        legend.text = element_text(size = 14),
        axis.title = element_text(size = 14)) + # +
  scale_fill_discrete(labels = c("men of color" = "Men of color (MOC)",
                        "white men" = "White men (WM)",
                        "white women" = "White women (WW)",
                        "women of color" = "Women of color (WOC)")) +
  labs(fill = "Participant race and gender") +
  scale_colour_discrete(guide="none") +
  theme(plot.margin = margin(.5,.1,0,.25, "cm"))  #(order goes top, right, bottom, left) 

```

## Studies 2a–2b: Mediator Studies
## Study 2a: Black Women and White Women
```{r s2dta, echo=FALSE, include = FALSE}

data = read_csv("~/Dropbox/Research/current projects/diversity gatekeeping/studies + data/lf mediator SP2025/march 2025 BW/data + analyses/lf mediator CLEAN data.csv")

data = data %>%
  mutate(race_name =
           case_when(prace == 1 ~ "Black women",
                     prace == 4 ~ "White women"
                     ))
table(data$race_name)

```

### Linked Fate
```{r s2lf1, echo=FALSE, include = FALSE}

lf_scale = subset(data, select = c(lfsb_1, lfsb_2, lfsb_3, lfsb_4, lfsb_5, lfsb_6))

lf_alpha = psych::alpha(lf_scale, check.keys = TRUE)
lf_alpha

data$linkedfate = furniture::rowmeans(data[, c("lfsb_1", "lfsb_2", "lfsb_3", "lfsb_4", "lfsb_5", "lfsb_6")])


summarySE(data, measurevar= c("linkedfate"), groupvars=c("race_name"), na.rm = TRUE, .drop = TRUE) 


lf_t = summarySE(data, measurevar= c("linkedfate"), groupvars=c("race_name"), na.rm = TRUE, .drop = TRUE) 

kable(lf_t,
      format = "html",
      align = 'c',
      booktabs = TRUE,
      digits = 2,
      caption = NA
      ) %>%
  kable_classic()

```

#### Figure
```{r s2lfp}
pd = position_dodge(.92)

lf_wb = data %>%
  ggplot(aes(x = race_name, y= linkedfate, fill = as.factor(race_name))) +
  geom_bar(stat = "summary", 
           fun = mean,
           position = "dodge") +
  stat_summary(fun.data=mean_se,
               geom='errorbar',
               #fun.args=list(conf.int=.95),
               linewidth=.6,
               aes(width=.3),
               color="black",
               position=pd) +
  coord_cartesian(ylim = c(1,7)) +
  scale_y_continuous(name = "Linked fate", expand = c(0,0), breaks=seq(1,7,1)) + # set ticks at 1, 2, 3, 4 +
    scale_x_discrete(name = "Comparison Group",
                   limits = c("Black women",
                              "White women"
                              )
                   #labels = c("linkedfate_woc" = "Linked fate with\nwomen of color",
                         #     "linkedfate_ww" = "Linked fate with\nWhite women")
                     ) +
 # scale_fill_discrete(labels = c("black_p" = "Black women", 
  #                             "white_p" = "White women")) +
  theme_classic() +
  theme(text = element_text(color = "black"),
        axis.text.x = element_text(size = 14, color = "black"),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size = 14, color = "black"),
        axis.title.y = element_text(size = 14, color = "black"),
        #axis.title.y = element_blank(),
        legend.title = element_text(size = 18, hjust = .5),
        legend.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        legend.position = "top",
        legend.title.position = "top"
        #legend.justification = "center",
        #legend.direction = "vertical",
        ) + 
  #labs(fill = "Participant race") +
  scale_colour_discrete(guide="none",
                        breaks = c("Black women" = "Black women", "White women" = "White women")
                        #labels = c("Black women" = "Black women", "White women" = "White women")
                        ) +
  scale_fill_manual(guide = "none",
                    name = "Participant Race", 
                    breaks = c("Black women" = "Black women", "White women" = "White women"),
                  #labels = c("black_p" = "Black women", "white_p" = "White women"),
                  values = c("#2FB3EA","#89939C")) +
  theme(plot.margin = margin(.5,.1,0,.25, "cm"))  #(order goes top, right, bottom, left) 

  
lf_wb
```

### Relational Orientation
 
```{r s2ro1, echo=FALSE, include = FALSE}
data$comp_wb_1r = 8 - data$comp_wb_1
data$comp_wb_3r = 8 - data$comp_wb_3
data$sup_wb_2r = 8 - data$sup_wb_2
data$sup_wb_3r = 8 - data$sup_wb_3
data$sup_wb_4r = 8 - data$sup_wb_4

supp_comp_wb_scale = subset(data, select = c(comp_wb_1r, comp_wb_2, comp_wb_3r, 
                                        sup_wb_1, sup_wb_2r, sup_wb_3r, sup_wb_4r))

supp_comp_wb_alpha = psych::alpha(supp_comp_wb_scale, check.keys = TRUE) 
supp_comp_wb_alpha

data$supp_comp_wb = furniture::rowmeans(data[, c("comp_wb_1r", "comp_wb_2", "comp_wb_3r", 
                                                 "sup_wb_1", "sup_wb_2r", "sup_wb_3r", "sup_wb_4r")])

summarySE(data = data, measurevar = "supp_comp_wb", groupvars = "race_name")
summarySE(data = data, measurevar = "supp_comp_wb")

```

```{r s2ro2, echo=FALSE, include = FALSE}
data$comp_bw_1r = 8 - data$comp_bw_1
data$comp_bw_3r = 8 - data$comp_bw_3
data$sup_bw_2r = 8 - data$sup_bw_2
data$sup_bw_3r = 8 - data$sup_bw_3
data$sup_bw_4r = 8 - data$sup_bw_4

supp_comp_bw_scale = subset(data, select = c(comp_bw_1r, comp_bw_2, comp_bw_3r, 
                                             sup_bw_1, sup_bw_2r, sup_bw_3r, sup_bw_4r))

supp_comp_bw_alpha = psych::alpha(supp_comp_bw_scale, check.keys = TRUE) 
supp_comp_bw_alpha
data$supp_comp_bw = furniture::rowmeans(data[, c("comp_bw_1r", "comp_bw_2", "comp_bw_3r", "sup_bw_1", "sup_bw_2r", "sup_bw_3r", "sup_bw_4r")])


summarySE(data = data, measurevar = "supp_comp_bw", groupvars = "race_name")
summarySE(data = data, measurevar = "supp_comp_bw")

```

#### Figures
##### White Women → Black Women
```{r s2ro3}
pd = position_dodge(.92)

ro_wb = data %>%
  ggplot(aes(x = race_name, y= supp_comp_wb, fill = as.factor(race_name))) +
  geom_bar(stat = "summary", 
           fun = mean,
           position = "dodge") +
  stat_summary(fun.data=mean_se,
               geom='errorbar',
               #fun.args=list(conf.int=.95),
               size=.6,
               aes(width=.3),
               color="black",
               position=pd) +
  coord_cartesian(ylim = c(1,7)) +
  scale_y_continuous(name = "Relational orientation", expand = c(0,0), breaks=seq(1,7,1)) + # set ticks at 1, 2, 3, 4 +
  scale_x_discrete(name = "Comparison Group",
                   limits = c("Black women",
                              "White women"
                              )
                   #labels = c("linkedfate_woc" = "Linked fate with\nwomen of color",
                         #     "linkedfate_ww" = "Linked fate with\nWhite women")
                     ) +
 # scale_fill_discrete(labels = c("black_p" = "Black women", 
  #                             "white_p" = "White women")) +
  theme_classic() +
  theme(text = element_text(color = "black"),
        axis.text.x = element_text(size = 14, color = "black"),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size = 14, color = "black"),
        axis.title.y = element_text(size = 14, color = "black"),
        #axis.title.y = element_blank(),
        legend.title = element_text(size = 18, hjust = .5),
        legend.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        legend.position = "top",
        legend.title.position = "top"
        #legend.justification = "center",
        #legend.direction = "vertical",
        ) + 
  #labs(fill = "Participant race") +
  scale_colour_discrete(guide="none",
                        breaks = c("Black women" = "Black women", "White women" = "White women")
                        #labels = c("Black women" = "Black women", "White women" = "White women")
                        ) +
  scale_fill_manual(guide = "none",
                    name = "Participant Race", 
                    breaks = c("Black women" = "Black women", "White women" = "White women"),
                  #labels = c("black_p" = "Black women", "white_p" = "White women"),
                  values = c("#2FB3EA","#89939C")) +
  theme(plot.margin = margin(.5,.1,0,.25, "cm"))  #(order goes top, right, bottom, left) 

ro_wb
```
##### Black Women → White Women
```{r s2ro4}
pd = position_dodge(.92)

ro_bw = data %>%
  ggplot(aes(x = race_name, y= supp_comp_bw, fill = as.factor(race_name))) +
  geom_bar(stat = "summary", 
           fun = mean,
           position = "dodge") +
  stat_summary(fun.data=mean_se,
               geom='errorbar',
               #fun.args=list(conf.int=.95),
               size=.6,
               aes(width=.3),
               color="black",
               position=pd) +
  coord_cartesian(ylim = c(1,7)) +
  scale_y_continuous(name = "Competition", expand = c(0,0), breaks=seq(1,7,1)) + # set ticks at 1, 2, 3, 4 +
  scale_x_discrete(name = "Comparison Group",
                   limits = c("Black women",
                              "White women"
                              )
                   #labels = c("linkedfate_woc" = "Linked fate with\nwomen of color",
                         #     "linkedfate_ww" = "Linked fate with\nWhite women")
                     ) +
 # scale_fill_discrete(labels = c("black_p" = "Black women", 
  #                             "white_p" = "White women")) +
  theme_classic() +
  theme(text = element_text(color = "black"),
        axis.text.x = element_text(size = 14, color = "black"),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size = 14, color = "black"),
        axis.title.y = element_blank(),
        legend.title = element_text(size = 18, hjust = .5),
        legend.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        legend.position = "top",
        legend.title.position = "top"
        #legend.justification = "center",
        #legend.direction = "vertical",
        ) + 
  #labs(fill = "Participant race") +
  scale_colour_discrete(guide="none",
                        breaks = c("Black women" = "Black women", "White women" = "White women")
                        #labels = c("Black women" = "Black women", "White women" = "White women")
                        ) +
  scale_fill_manual(guide = "none",
                    name = "Participant Race", 
                    breaks = c("Black women" = "Black women", "White women" = "White women"),
                  #labels = c("black_p" = "Black women", "white_p" = "White women"),
                  values = c("#2FB3EA","#89939C")) +
  theme(plot.margin = margin(.5,.1,0,.25, "cm"))  #(order goes top, right, bottom, left) 


ro_bw
```

### Bias Map

#### Pity Elicitation

```{r s2e1, echo=FALSE, include = FALSE}

e_ww_scale = subset(data, select = c(bm_ww_e_1, bm_ww_e_2))

e_ww_corr = corr.test(e_ww_scale, method = "pearson")
e_ww_corr 

library(corrr)
e_ww_corr_t = e_ww_scale %>%
  correlate() %>%
  shave(upper = TRUE) %>%
  fashion(decimals = 2, na_print = " ") 

kable(e_ww_corr_t,
      format = "html",
      align = 'c',
      booktabs = TRUE,
      digits = 2,
      caption = NA
      ) %>%
  kable_classic()

data$emotion_ww = furniture::rowmeans(data[, c("bm_ww_e_1", "bm_ww_e_2")])

summarySE(data, measurevar= c("emotion_ww"), groupvars=c("race_name"), na.rm = TRUE) 

bm_ww_e_t = summarySE(data, measurevar= c("emotion_ww"), groupvars=c("race_name"), na.rm = TRUE, .drop = TRUE) 

kable(bm_ww_e_t,
      format = "html",
      align = 'c',
      booktabs = TRUE,
      digits = 2,
      caption = NA
      ) %>%
  kable_classic()

```

```{r s2e2, echo=FALSE, include = FALSE}
e_bw_scale = subset(data, select = c(bm_bw_e_1, bm_bw_e_2))

e_bw_corr = corr.test(e_bw_scale, method = "pearson")
e_bw_corr 

library(corrr)
e_bw_corr_t = e_bw_scale %>%
  correlate() %>%
  shave(upper = TRUE) %>%
  fashion(decimals = 2, na_print = " ") 

kable(e_bw_corr_t,
      format = "html",
      align = 'c',
      booktabs = TRUE,
      digits = 2,
      caption = NA
      ) %>%
  kable_classic()

data$emotion_bw = furniture::rowmeans(data[, c("bm_bw_e_1", "bm_bw_e_2")])

summarySE(data, measurevar= c("emotion_bw"), groupvars=c("race_name"), na.rm = TRUE) 

bm_bw_e_t = summarySE(data, measurevar= c("emotion_bw"), groupvars=c("race_name"), na.rm = TRUE, .drop = TRUE) 

kable(bm_bw_e_t,
      format = "html",
      align = 'c',
      booktabs = TRUE,
      digits = 2,
      caption = NA
      ) %>%
  kable_classic()
```

#### Figures
##### White Women → Black Women
```{r s2e3}
pd = position_dodge(.92)

e_wb = data %>%
  ggplot(aes(x = race_name, y= emotion_ww, fill = as.factor(race_name))) +
  geom_bar(stat = "summary", 
           fun = mean,
           position = "dodge") +
  stat_summary(fun.data=mean_se,
               geom='errorbar',
               #fun.args=list(conf.int=.95),
               size=.6,
               aes(width=.3),
               color="black",
               position=pd) +
  coord_cartesian(ylim = c(1,7)) +
  scale_y_continuous(name = "Pity elicitation", expand = c(0,0), breaks=seq(1,7,1)) + # set ticks at 1, 2, 3, 4 +
  scale_x_discrete(name = "Comparison Group",
                   limits = c("Black women",
                              "White women"
                              )
                   #labels = c("linkedfate_woc" = "Linked fate with\nwomen of color",
                         #     "linkedfate_ww" = "Linked fate with\nWhite women")
                     ) +
 # scale_fill_discrete(labels = c("black_p" = "Black women", 
  #                             "white_p" = "White women")) +
  theme_classic() +
  theme(text = element_text(color = "black"),
        axis.text.x = element_text(size = 14, color = "black"),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size = 14, color = "black"),
        axis.title.y = element_text(size = 14, color = "black"),
        #axis.title.y = element_blank(),
        legend.title = element_text(size = 18, hjust = .5),
        legend.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        legend.position = "top",
        legend.title.position = "top"
        #legend.justification = "center",
        #legend.direction = "vertical",
        ) + 
  #labs(fill = "Participant race") +
  scale_colour_discrete(guide="none",
                        breaks = c("Black women" = "Black women", "White women" = "White women")
                        #labels = c("Black women" = "Black women", "White women" = "White women")
                        ) +
  scale_fill_manual(guide = "none",
                    name = "Participant Race", 
                    breaks = c("Black women" = "Black women", "White women" = "White women"),
                  #labels = c("black_p" = "Black women", "white_p" = "White women"),
                  values = c("#2FB3EA","#89939C")) +
  theme(plot.margin = margin(.5,.1,0,.25, "cm"))  #(order goes top, right, bottom, left) 

e_wb
```

##### Black Women → White Women

```{r s2e4}
pd = position_dodge(.92)

e_bw = data %>%
  ggplot(aes(x = race_name, y= emotion_bw, fill = as.factor(race_name))) +
  geom_bar(stat = "summary", 
           fun = mean,
           position = "dodge") +
  stat_summary(fun.data=mean_se,
               geom='errorbar',
               #fun.args=list(conf.int=.95),
               size=.6,
               aes(width=.3),
               color="black",
               position=pd) +
  coord_cartesian(ylim = c(1,7)) +
  scale_y_continuous(name = "Pity elicitation", expand = c(0,0), breaks=seq(1,7,1)) + # set ticks at 1, 2, 3, 4 +
  scale_x_discrete(name = "Comparison Group",
                   limits = c("Black women",
                              "White women"
                              )
                   #labels = c("linkedfate_woc" = "Linked fate with\nwomen of color",
                         #     "linkedfate_ww" = "Linked fate with\nWhite women")
                     ) +
 # scale_fill_discrete(labels = c("black_p" = "Black women", 
  #                             "white_p" = "White women")) +
  theme_classic() +
  theme(text = element_text(color = "black"),
        axis.text.x = element_text(size = 14, color = "black"),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size = 14, color = "black"),
        axis.title.y = element_blank(),
        legend.title = element_text(size = 18, hjust = .5),
        legend.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        legend.position = "top",
        legend.title.position = "top"
        #legend.justification = "center",
        #legend.direction = "vertical",
        ) + 
  #labs(fill = "Participant race") +
  scale_colour_discrete(guide="none",
                        breaks = c("Black women" = "Black women", "White women" = "White women")
                        #labels = c("Black women" = "Black women", "White women" = "White women")
                        ) +
  scale_fill_manual(guide = "none",
                    name = "Participant Race", 
                    breaks = c("Black women" = "Black women", "White women" = "White women"),
                  #labels = c("black_p" = "Black women", "white_p" = "White women"),
                  values = c("#2FB3EA","#89939C")) +
  theme(plot.margin = margin(.5,.1,0,.25, "cm"))  #(order goes top, right, bottom, left) 

e_bw
```

#### Active Facilitation

```{r s2b1, echo=FALSE, include = FALSE}

bm_ww_b_scale = subset(data, select = c(bm_ww_b_1, bm_ww_b_2, bm_ww_b_3, bm_ww_b_4))

bm_ww_b_alpha = psych::alpha(bm_ww_b_scale, check.keys = TRUE) #alpha = .
bm_ww_b_alpha

data$behavior_ww = furniture::rowmeans(data[, c("bm_ww_b_1", "bm_ww_b_2", "bm_ww_b_3", "bm_ww_b_4")])

summarySE(data, measurevar= c("behavior_ww"), groupvars=c("race_name"), na.rm = TRUE) 

bm_ww_b_t = summarySE(data, measurevar= c("behavior_ww"), groupvars=c("race_name"), na.rm = TRUE, .drop = TRUE) 

kable(bm_ww_b_t,
      format = "html",
      align = 'c',
      booktabs = TRUE,
      digits = 2,
      caption = NA
      ) %>%
  kable_classic()

```

```{r s2b2, echo=FALSE, include = FALSE}
bm_bw_b_scale = subset(data, select = c(bm_bw_b_1, bm_bw_b_2, bm_bw_b_3, bm_bw_b_4))

bm_bw_b_alpha = psych::alpha(bm_bw_b_scale, check.keys = TRUE) #alpha = .
bm_bw_b_alpha

data$behavior_bw = furniture::rowmeans(data[, c("bm_bw_b_1", "bm_bw_b_2", "bm_bw_b_3", "bm_bw_b_4")])

summarySE(data, measurevar= c("behavior_bw"), groupvars=c("race_name"), na.rm = TRUE) 

bm_bw_b_t = summarySE(data, measurevar= c("behavior_bw"), groupvars=c("race_name"), na.rm = TRUE, .drop = TRUE) 

kable(bm_bw_b_t,
      format = "html",
      align = 'c',
      booktabs = TRUE,
      digits = 2,
      caption = NA
      ) %>%
  kable_classic()
```

#### Figures

##### White Women → Black Women
```{r s2b3}
pd = position_dodge(.92)

b_wb = data %>%
  ggplot(aes(x = race_name, y= behavior_ww, fill = as.factor(race_name))) +
  geom_bar(stat = "summary", 
           fun = mean,
           position = "dodge") +
  stat_summary(fun.data=mean_se,
               geom='errorbar',
               #fun.args=list(conf.int=.95),
               size=.6,
               aes(width=.3),
               color="black",
               position=pd) +
  coord_cartesian(ylim = c(1,7)) +
  scale_y_continuous(name = "Active facilitation", expand = c(0,0), breaks=seq(1,7,1)) + # set ticks at 1, 2, 3, 4 +
   scale_x_discrete(name = "Comparison Group",
                   limits = c("Black women",
                              "White women"
                              )
                   #labels = c("linkedfate_woc" = "Linked fate with\nwomen of color",
                         #     "linkedfate_ww" = "Linked fate with\nWhite women")
                     ) +
 # scale_fill_discrete(labels = c("black_p" = "Black women", 
  #                             "white_p" = "White women")) +
  theme_classic() +
  theme(text = element_text(color = "black"),
        axis.text.x = element_text(size = 14, color = "black"),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size = 14, color = "black"),
        axis.title.y = element_text(size = 14, color = "black"),
       # axis.title.y = element_blank(),
        legend.title = element_text(size = 18, hjust = .5),
        legend.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        legend.position = "top",
        legend.title.position = "top"
        #legend.justification = "center",
        #legend.direction = "vertical",
        ) + 
  #labs(fill = "Participant race") +
  scale_colour_discrete(guide="none",
                        breaks = c("Black women" = "Black women", "White women" = "White women")
                        #labels = c("Black women" = "Black women", "White women" = "White women")
                        ) +
  scale_fill_manual(guide="none",
                    name = "Participant Race", 
                    breaks = c("Black women" = "Black women", "White women" = "White women"),
                  #labels = c("black_p" = "Black women", "white_p" = "White women"),
                  values = c("#2FB3EA","#89939C")) +
  theme(plot.margin = margin(.5,.1,0,.25, "cm"))  #(order goes top, right, bottom, left) 

b_wb
```

##### Black Women → White Women

```{r s2b4}
pd = position_dodge(.92)

b_bw = data %>%
  ggplot(aes(x = race_name, y= behavior_bw, fill = as.factor(race_name))) +
  geom_bar(stat = "summary", 
           fun = mean,
           position = "dodge") +
  stat_summary(fun.data=mean_se,
               geom='errorbar',
               #fun.args=list(conf.int=.95),
               size=.6,
               aes(width=.3),
               color="black",
               position=pd) +
  coord_cartesian(ylim = c(1,7)) +
  scale_y_continuous(name = "Active faciliation", expand = c(0,0), breaks=seq(1,7,1)) + # set ticks at 1, 2, 3, 4 +
  scale_x_discrete(name = "Comparison Group",
                   limits = c("Black women",
                              "White women"
                              )
                   #labels = c("linkedfate_woc" = "Linked fate with\nwomen of color",
                         #     "linkedfate_ww" = "Linked fate with\nWhite women")
                     ) +
 # scale_fill_discrete(labels = c("black_p" = "Black women", 
  #                             "white_p" = "White women")) +
  theme_classic() +
  theme(text = element_text(color = "black"),
        axis.text.x = element_text(size = 14, color = "black"),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size = 14, color = "black"),
        axis.title.y = element_blank(),
        legend.title = element_text(size = 18, hjust = .5),
        legend.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        legend.position = "top",
        legend.title.position = "top"
        #legend.justification = "center",
        #legend.direction = "vertical",
        ) + 
  #labs(fill = "Participant race") +
  scale_colour_discrete(guide="none",
                        breaks = c("Black women" = "Black women", "White women" = "White women")
                        #labels = c("Black women" = "Black women", "White women" = "White women")
                        ) +
  scale_fill_manual(guide="none",
                    name = "Participant Race", 
                    breaks = c("Black women" = "Black women", "White women" = "White women"),
                  #labels = c("black_p" = "Black women", "white_p" = "White women"),
                  values = c("#2FB3EA","#89939C")) +
  theme(plot.margin = margin(.5,.1,0,.25, "cm"))  #(order goes top, right, bottom, left) 


b_bw
```

### Study 2a Combined Figure

```{r s2b6}
ggarrange(ro_wb, ro_bw,
          e_wb, e_bw,
          b_wb, b_bw,
          nrow = 3,
          ncol = 2)
```

## Study 2b: Latina Women and White Women

```{r s2bdta, echo=FALSE, include = FALSE}
data = read_csv("~/Dropbox/Research/current projects/diversity gatekeeping/studies + data/lf mediator SP2025/march 2025 LW/data + analyses/lf mediator Mar25 LW clean.csv")

data = data %>%
  mutate(race_name =
           case_when(prace == 2 ~ "Latina women",
                     prace == 4 ~ "White women"
                     ))
table(data$race_name)

```

### Linked Fate
```{r s2blf1, echo=FALSE, include = FALSE}

lf_scale = subset(data, select = c(lfsb_1, lfsb_2, lfsb_3, lfsb_4, lfsb_5, lfsb_6))

lf_alpha = psych::alpha(lf_scale, check.keys = TRUE)
lf_alpha

data$linkedfate = furniture::rowmeans(data[, c("lfsb_1", "lfsb_2", "lfsb_3", "lfsb_4", "lfsb_5", "lfsb_6")])


summarySE(data, measurevar= c("linkedfate"), groupvars=c("race_name"), na.rm = TRUE, .drop = TRUE) 


lf_t = summarySE(data, measurevar= c("linkedfate"), groupvars=c("race_name"), na.rm = TRUE, .drop = TRUE) 

kable(lf_t,
      format = "html",
      align = 'c',
      booktabs = TRUE,
      digits = 2,
      caption = NA
      ) %>%
  kable_classic()

```

#### Figure
```{r s2blf2}
pd = position_dodge(.92)

lf_wl = data %>%
  ggplot(aes(x = race_name, y= linkedfate, fill = as.factor(race_name))) +
  geom_bar(stat = "summary", 
           fun = mean,
           position = "dodge") +
  stat_summary(fun.data=mean_se,
               geom='errorbar',
               #fun.args=list(conf.int=.95),
               size=.6,
               aes(width=.3),
               color="black",
               position=pd) +
  coord_cartesian(ylim = c(1,7)) +
  scale_y_continuous(name = "Linked fate", expand = c(0,0), breaks=seq(1,7,1)) + # set ticks at 1, 2, 3, 4
    scale_x_discrete(name = "Comparison Group",
                   limits = c("Latina women",
                              "White women"
                              )
                   #labels = c("linkedfate_woc" = "Linked fate with\nwomen of color",
                         #     "linkedfate_ww" = "Linked fate with\nWhite women")
                     ) +
 # scale_fill_discrete(labels = c("black_p" = "Black women", 
  #                             "white_p" = "White women")) +
  theme_classic() +
  theme(text = element_text(color = "black"),
        axis.text.x = element_text(size = 14, color = "black"),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size = 14, color = "black"),
        axis.title.y = element_text(size = 14, color = "black"),
        #axis.title.y = element_blank(),
        legend.title = element_text(size = 18, hjust = .5),
        legend.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        legend.position = "top",
        legend.title.position = "top"
        #legend.justification = "center",
        #legend.direction = "vertical",
        ) + 
  #labs(fill = "Participant race") +
  scale_colour_discrete(guide="none",
                        breaks = c("Latina women" = "Latina women","White women" = "White women")
                        #labels = c("Latina women" = "Latina women", "White women" = "White women")
                        ) +
  scale_fill_manual(guide="none",
                    name = "Participant Race", 
                    breaks = c("Latina women" = "Latina women", "White women" = "White women"),
                  #labels = c("latina_p" = "Latina women", "white_p" = "White women"),
                  values = c("#0078AB", "#89939C")) +
  theme(plot.margin = margin(.5,.1,0,.25, "cm"))  #(order goes top, right, bottom, left) 

lf_wl
```

### Relational Orientation

```{r s2bro1, echo=FALSE, include = FALSE}
data$comp_wl_1r = 8 - data$comp_wl_1
data$comp_wl_3r = 8 - data$comp_wl_3
data$sup_wl_2r = 8 - data$sup_wl_2
data$sup_wl_3r = 8 - data$sup_wl_3
data$sup_wl_4r = 8 - data$sup_wl_4

supp_comp_wl_scale = subset(data, select = c(comp_wl_1r, comp_wl_2, comp_wl_3r, 
                                        sup_wl_1, sup_wl_2r, sup_wl_3r, sup_wl_4r))

supp_comp_wl_alpha = psych::alpha(supp_comp_wl_scale, check.keys = TRUE) 
supp_comp_wl_alpha

data$supp_comp_wl = furniture::rowmeans(data[, c("comp_wl_1r", "comp_wl_2", "comp_wl_3r", 
                                                 "sup_wl_1", "sup_wl_2r", "sup_wl_3r", "sup_wl_4r")])

summarySE(data = data, measurevar = "supp_comp_wl", groupvars = "race_name")
summarySE(data = data, measurevar = "supp_comp_wl")

```

#### Latina Women → White Women

```{r s2bro2, echo=FALSE, include = FALSE}
data$comp_lw_1r = 8 - data$comp_lw_1
data$comp_lw_3r = 8 - data$comp_lw_3
data$sup_lw_2r = 8 - data$sup_lw_2
data$sup_lw_3r = 8 - data$sup_lw_3
data$sup_lw_4r = 8 - data$sup_lw_4

supp_comp_lw_scale = subset(data, select = c(comp_lw_1r, comp_lw_2, comp_lw_3r, 
                                             sup_lw_1, sup_lw_2r, sup_lw_3r, sup_lw_4r))

supp_comp_lw_alpha = psych::alpha(supp_comp_lw_scale, check.keys = TRUE) 
supp_comp_lw_alpha
data$supp_comp_lw = furniture::rowmeans(data[, c("comp_lw_1r", "comp_lw_2", "comp_lw_3r", "sup_lw_1", "sup_lw_2r", "sup_lw_3r", "sup_lw_4r")])


summarySE(data = data, measurevar = "supp_comp_lw", groupvars = "race_name")
summarySE(data = data, measurevar = "supp_comp_lw")

```

#### Figures
##### White Women → Latina Women
```{r s2bro3}
pd = position_dodge(.92)

ro_wl = data %>%
  ggplot(aes(x = race_name, y= supp_comp_wl, fill = as.factor(race_name))) +
  geom_bar(stat = "summary", 
           fun = mean,
           position = "dodge") +
  stat_summary(fun.data=mean_se,
               geom='errorbar',
               #fun.args=list(conf.int=.95),
               size=.6,
               aes(width=.3),
               color="black",
               position=pd) +
  coord_cartesian(ylim = c(1,7)) +
  scale_y_continuous(name = "Relational orientation", expand = c(0,0), breaks=seq(1,7,1)) + # set ticks at 1, 2, 3, 4 +
  scale_x_discrete(name = "Comparison Group",
                   limits = c("Latina women",
                              "White women"
                              )
                   #labels = c("linkedfate_woc" = "Linked fate with\nwomen of color",
                         #     "linkedfate_ww" = "Linked fate with\nWhite women")
                     ) +
 # scale_fill_discrete(labels = c("black_p" = "Black women", 
  #                             "white_p" = "White women")) +
  theme_classic() +
  theme(text = element_text(color = "black"),
        axis.text.x = element_text(size = 14, color = "black"),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size = 14, color = "black"),
        axis.title.y = element_text(size = 14, color = "black"),
        #axis.title.y = element_blank(),
        legend.title = element_text(size = 18, hjust = .5),
        legend.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        legend.position = "top",
        legend.title.position = "top"
        #legend.justification = "center",
        #legend.direction = "vertical",
        ) + 
  #labs(fill = "Participant race") +
  scale_colour_discrete(guide="none",
                        breaks = c("Latina women" = "Latina women","White women" = "White women")
                        #labels = c("Latina women" = "Latina women", "White women" = "White women")
                        ) +
  scale_fill_manual(guide="none",
                    name = "Participant Race", 
                    breaks = c("Latina women" = "Latina women", "White women" = "White women"),
                  #labels = c("latina_p" = "Latina women", "white_p" = "White women"),
                  values = c("#0078AB", "#89939C")) +
  theme(plot.margin = margin(.5,.1,0,.25, "cm"))  #(order goes top, right, bottom, left) 

ro_wl
```

##### Latina Women → White Women

```{r s2bro4}
pd = position_dodge(.92)

ro_lw = data %>%
  ggplot(aes(x = race_name, y= supp_comp_lw, fill = as.factor(race_name))) +
  geom_bar(stat = "summary", 
           fun = mean,
           position = "dodge") +
  stat_summary(fun.data=mean_se,
               geom='errorbar',
               #fun.args=list(conf.int=.95),
               size=.6,
               aes(width=.3),
               color="black",
               position=pd) +
  coord_cartesian(ylim = c(1,7)) +
  scale_y_continuous(name = "Relational orientation", expand = c(0,0), breaks=seq(1,7,1)) + # set ticks at 1, 2, 3, 4 +
  scale_x_discrete(name = "Comparison Group",
                   limits = c("Latina women",
                              "White women"
                              )
                   #labels = c("linkedfate_woc" = "Linked fate with\nwomen of color",
                         #     "linkedfate_ww" = "Linked fate with\nWhite women")
                     ) +
 # scale_fill_discrete(labels = c("black_p" = "Black women", 
  #                             "white_p" = "White women")) +
  theme_classic() +
  theme(text = element_text(color = "black"),
        axis.text.x = element_text(size = 14, color = "black"),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size = 14, color = "black"),
        axis.title.y = element_blank(),
        legend.title = element_text(size = 18, hjust = .5),
        legend.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        legend.position = "top",
        legend.title.position = "top"
        #legend.justification = "center",
        #legend.direction = "vertical",
        ) + 
  #labs(fill = "Participant race") +
  scale_colour_discrete(guide="none",
                        breaks = c("Latina women" = "Latina women","White women" = "White women")
                        #labels = c("Latina women" = "Latina women", "White women" = "White women")
                        ) +
  scale_fill_manual(guide="none",
                    name = "Participant Race", 
                    breaks = c("Latina women" = "Latina women", "White women" = "White women"),
                  #labels = c("latina_p" = "Latina women", "white_p" = "White women"),
                  values = c("#0078AB", "#89939C")) +
  theme(plot.margin = margin(.5,.1,0,.25, "cm"))  #(order goes top, right, bottom, left) 

ro_lw
```

### Bias Map

#### Pity Elicitation

```{r s2be1, echo=FALSE, include = FALSE}

e_ww_scale = subset(data, select = c(bm_ww_e_1, bm_ww_e_2))

e_ww_corr = corr.test(e_ww_scale, method = "pearson")
e_ww_corr 

library(corrr)
e_ww_corr_t = e_ww_scale %>%
  correlate() %>%
  shave(upper = TRUE) %>%
  fashion(decimals = 2, na_print = " ") 

kable(e_ww_corr_t,
      format = "html",
      align = 'c',
      booktabs = TRUE,
      digits = 2,
      caption = NA
      ) %>%
  kable_classic()

data$emotion_ww = furniture::rowmeans(data[, c("bm_ww_e_1", "bm_ww_e_2")])

summarySE(data, measurevar= c("emotion_ww"), groupvars=c("race_name"), na.rm = TRUE) 

bm_ww_e_t = summarySE(data, measurevar= c("emotion_ww"), groupvars=c("race_name"), na.rm = TRUE, .drop = TRUE) 

kable(bm_ww_e_t,
      format = "html",
      align = 'c',
      booktabs = TRUE,
      digits = 2,
      caption = NA
      ) %>%
  kable_classic()

```

```{r s2be2, echo=FALSE, include = FALSE}
e_lw_scale = subset(data, select = c(bm_lw_e_1, bm_lw_e_2))

e_lw_corr = corr.test(e_lw_scale, method = "pearson")
e_lw_corr 

library(corrr)
e_lw_corr_t = e_lw_scale %>%
  correlate() %>%
  shave(upper = TRUE) %>%
  fashion(decimals = 2, na_print = " ") 

kable(e_lw_corr_t,
      format = "html",
      align = 'c',
      booktabs = TRUE,
      digits = 2,
      caption = NA
      ) %>%
  kable_classic()

data$emotion_lw = furniture::rowmeans(data[, c("bm_lw_e_1", "bm_lw_e_2")])

summarySE(data, measurevar= c("emotion_lw"), groupvars=c("race_name"), na.rm = TRUE) 

bm_lw_e_t = summarySE(data, measurevar= c("emotion_lw"), groupvars=c("race_name"), na.rm = TRUE, .drop = TRUE) 

kable(bm_lw_e_t,
      format = "html",
      align = 'c',
      booktabs = TRUE,
      digits = 2,
      caption = NA
      ) %>%
  kable_classic()
```

#### Figures

##### White Women → Latina Women

```{r s2be3}
pd = position_dodge(.92)

e_wl = data %>%
  ggplot(aes(x = race_name, y= emotion_ww, fill = as.factor(race_name))) +
  geom_bar(stat = "summary", 
           fun = mean,
           position = "dodge") +
  stat_summary(fun.data=mean_se,
               geom='errorbar',
               #fun.args=list(conf.int=.95),
               size=.6,
               aes(width=.3),
               color="black",
               position=pd) +
  coord_cartesian(ylim = c(1,7)) +
  scale_y_continuous(name = "Pity elicitation", expand = c(0,0), breaks=seq(1,7,1)) + # set ticks at 1, 2, 3, 4 +
  scale_x_discrete(name = "Comparison Group",
                   limits = c("Latina women",
                              "White women"
                              )
                   #labels = c("linkedfate_woc" = "Linked fate with\nwomen of color",
                         #     "linkedfate_ww" = "Linked fate with\nWhite women")
                     ) +
 # scale_fill_discrete(labels = c("black_p" = "Black women", 
  #                             "white_p" = "White women")) +
  theme_classic() +
  theme(text = element_text(color = "black"),
        axis.text.x = element_text(size = 14, color = "black"),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size = 14, color = "black"),
        axis.title.y = element_text(size = 14, color = "black"),
        #axis.title.y = element_blank(),
        legend.title = element_text(size = 18, hjust = .5),
        legend.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        legend.position = "top",
        legend.title.position = "top"
        #legend.justification = "center",
        #legend.direction = "vertical",
        ) + 
  #labs(fill = "Participant race") +
  scale_colour_discrete(guide="none",
                        breaks = c("Latina women" = "Latina women","White women" = "White women")
                        #labels = c("Latina women" = "Latina women", "White women" = "White women")
                        ) +
  scale_fill_manual(guide="none",
                    name = "Participant Race", 
                    breaks = c("Latina women" = "Latina women", "White women" = "White women"),
                  #labels = c("latina_p" = "Latina women", "white_p" = "White women"),
                  values = c("#0078AB", "#89939C")) +
  theme(plot.margin = margin(.5,.1,0,.25, "cm"))  #(order goes top, right, bottom, left) 

e_wl
```

##### Latina Women → White Women

```{r s2be4}
pd = position_dodge(.92)

e_lw = data %>%
  ggplot(aes(x = race_name, y= emotion_lw, fill = as.factor(race_name))) +
  geom_bar(stat = "summary", 
           fun = mean,
           position = "dodge") +
  stat_summary(fun.data=mean_se,
               geom='errorbar',
               #fun.args=list(conf.int=.95),
               size=.6,
               aes(width=.3),
               color="black",
               position=pd) +
  coord_cartesian(ylim = c(1,7)) +
  scale_y_continuous(name = "Pity elicitation", expand = c(0,0), breaks=seq(1,7,1)) + # set ticks at 1, 2, 3, 4 +
   scale_x_discrete(name = "Comparison Group",
                   limits = c("Latina women",
                              "White women"
                              )
                   #labels = c("linkedfate_woc" = "Linked fate with\nwomen of color",
                         #     "linkedfate_ww" = "Linked fate with\nWhite women")
                     ) +
 # scale_fill_discrete(labels = c("black_p" = "Black women", 
  #                             "white_p" = "White women")) +
  theme_classic() +
  theme(text = element_text(color = "black"),
        axis.text.x = element_text(size = 14, color = "black"),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size = 14, color = "black"),
        axis.title.y = element_blank(),
        legend.title = element_text(size = 18, hjust = .5),
        legend.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        legend.position = "top",
        legend.title.position = "top"
        #legend.justification = "center",
        #legend.direction = "vertical",
        ) + 
  #labs(fill = "Participant race") +
  scale_colour_discrete(guide="none",
                        breaks = c("Latina women" = "Latina women","White women" = "White women")
                        #labels = c("Latina women" = "Latina women", "White women" = "White women")
                        ) +
  scale_fill_manual(guide="none",
                    name = "Participant Race", 
                    breaks = c("Latina women" = "Latina women", "White women" = "White women"),
                  #labels = c("latina_p" = "Latina women", "white_p" = "White women"),
                  values = c("#0078AB", "#89939C")) +
  theme(plot.margin = margin(.5,.1,0,.25, "cm"))  #(order goes top, right, bottom, left) 

e_lw
```

#### Active Facilitation

```{r s2bb1, echo=FALSE, include = FALSE}

bm_ww_b_scale = subset(data, select = c(bm_ww_b_1, bm_ww_b_2, bm_ww_b_3, bm_ww_b_4))

bm_ww_b_alpha = psych::alpha(bm_ww_b_scale, check.keys = TRUE) #alpha = .
bm_ww_b_alpha

data$behavior_ww = furniture::rowmeans(data[, c("bm_ww_b_1", "bm_ww_b_2", "bm_ww_b_3", "bm_ww_b_4")])

summarySE(data, measurevar= c("behavior_ww"), groupvars=c("race_name"), na.rm = TRUE) 

bm_ww_b_t = summarySE(data, measurevar= c("behavior_ww"), groupvars=c("race_name"), na.rm = TRUE, .drop = TRUE) 

kable(bm_ww_b_t,
      format = "html",
      align = 'c',
      booktabs = TRUE,
      digits = 2,
      caption = NA
      ) %>%
  kable_classic()
```


```{r s2bb2, echo=FALSE, include = FALSE}
bm_lw_b_scale = subset(data, select = c(bm_lw_b_1, bm_lw_b_2, bm_lw_b_3, bm_lw_b_4))

bm_lw_b_alpha = psych::alpha(bm_lw_b_scale, check.keys = TRUE) #alpha = .
bm_lw_b_alpha

data$behavior_lw = furniture::rowmeans(data[, c("bm_lw_b_1", "bm_lw_b_2", "bm_lw_b_3", "bm_lw_b_4")])

summarySE(data, measurevar= c("behavior_lw"), groupvars=c("race_name"), na.rm = TRUE) 

bm_lw_b_t = summarySE(data, measurevar= c("behavior_lw"), groupvars=c("race_name"), na.rm = TRUE, .drop = TRUE) 

kable(bm_lw_b_t,
      format = "html",
      align = 'c',
      booktabs = TRUE,
      digits = 2,
      caption = NA
      ) %>%
  kable_classic()


```

#### Figures

##### White Women → Latina Women
```{r s2bb3}
pd = position_dodge(.92)

b_wl = data %>%
  ggplot(aes(x = race_name, y= behavior_ww, fill = as.factor(race_name))) +
  geom_bar(stat = "summary", 
           fun = mean,
           position = "dodge") +
  stat_summary(fun.data=mean_se,
               geom='errorbar',
               #fun.args=list(conf.int=.95),
               size=.6,
               aes(width=.3),
               color="black",
               position=pd) +
  coord_cartesian(ylim = c(1,7)) +
  scale_y_continuous(name = "Active facilitation", expand = c(0,0), breaks=seq(1,7,1)) + # set ticks at 1, 2, 3, 4 +
  scale_x_discrete(name = "Comparison Group",
                   limits = c("Latina women",
                              "White women"
                              )
                   #labels = c("linkedfate_woc" = "Linked fate with\nwomen of color",
                         #     "linkedfate_ww" = "Linked fate with\nWhite women")
                     ) +
 # scale_fill_discrete(labels = c("black_p" = "Black women", 
  #                             "white_p" = "White women")) +
  theme_classic() +
  theme(text = element_text(color = "black"),
        axis.text.x = element_text(size = 14, color = "black"),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size = 14, color = "black"),
        axis.title.y = element_text(size = 14, color = "black"),
        #axis.title.y = element_blank(),
        legend.title = element_text(size = 18, hjust = .5),
        legend.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        legend.position = "top",
        legend.title.position = "top"
        #legend.justification = "center",
        #legend.direction = "vertical",
        ) + 
  #labs(fill = "Participant race") +
  scale_colour_discrete(guide="none",
                        breaks = c("Latina women" = "Latina women","White women" = "White women")
                        #labels = c("Latina women" = "Latina women", "White women" = "White women")
                        ) +
  scale_fill_manual(guide="none",
                    name = "Participant Race", 
                    breaks = c("Latina women" = "Latina women", "White women" = "White women"),
                  #labels = c("latina_p" = "Latina women", "white_p" = "White women"),
                  values = c("#0078AB", "#89939C")) +
  theme(plot.margin = margin(.5,.1,0,.25, "cm"))  #(order goes top, right, bottom, left) 

b_wl
```

##### Latina Women → White Women

```{r s2bb4}
pd = position_dodge(.92)

b_lw = data %>%
  ggplot(aes(x = race_name, y= behavior_lw, fill = as.factor(race_name))) +
  geom_bar(stat = "summary", 
           fun = mean,
           position = "dodge") +
  stat_summary(fun.data=mean_se,
               geom='errorbar',
               #fun.args=list(conf.int=.95),
               size=.6,
               aes(width=.3),
               color="black",
               position=pd) +
  coord_cartesian(ylim = c(1,7)) +
  scale_y_continuous(name = "Active facilitation", expand = c(0,0), breaks=seq(1,7,1)) + # set ticks at 1, 2, 3, 4 +
  scale_x_discrete(name = "Comparison Group",
                   limits = c("White women",
                              "Latina women")
                   #labels = c("linkedfate_woc" = "Linked fate with\nwomen of color",
                         #     "linkedfate_ww" = "Linked fate with\nWhite women")
                     ) +
 # scale_fill_discrete(labels = c("black_p" = "Black women", 
  #                             "white_p" = "White women")) +
  theme_classic() +
  theme(text = element_text(color = "black"),
        axis.text.x = element_text(size = 14, color = "black"),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size = 14, color = "black"),
        axis.title.y = element_blank(),
        legend.title = element_text(size = 18, hjust = .5),
        legend.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        legend.position = "top",
        legend.title.position = "top"
        #legend.justification = "center",
        #legend.direction = "vertical",
        ) + # +
  #labs(fill = "Participant race") +
  scale_colour_discrete(guide="none",
                        breaks = c("White women" = "White women", "Latina women" = "Latina women")#,
                        #labels = c("White women" = "White women", "Latina women" = "Latina women")
                        ) +
  scale_fill_manual(guide="none",
                    name = "Participant Race", 
                  breaks= c("White women" = "White women", "Latina women" = "Latina women"),
                 # labels = c("white_p" = "White women", "latina_p" = "Latina women") #,
                  values = c("#0078AB", "#89939C")) +
  theme(plot.margin = margin(.5,.1,0,.25, "cm"))  #(order goes top, right, bottom, left) 

b_lw
```

### Study 2b Combined Figure

```{r s2all1}

ggarrange(ro_wl, ro_lw,
          e_wl, e_lw,
          b_wl, b_lw,
          nrow = 3,
          ncol = 2)

# Study 2b plot export
```

## Study 3a: Black Women and White Women

```{r s3adta, echo=FALSE, include = FALSE}

data_bw <- read_csv("~/Dropbox/Research/current projects/diversity gatekeeping/studies + data/workshop study Oct:Nov 24/data + analyses/bw_ww/gen div workshop BW_WW Oct 24 CLEAN.csv")

```

### Self-Applicability

```{r s3aselfapplicability}

pd = position_dodge(.92)

bsa = data_bw %>%
  ggplot(aes(x = webad_race, y= sa_mean, fill = prace_name)) +
  geom_bar(stat = "summary", 
           fun = mean,
           position = "dodge") +
  stat_summary(fun.data=mean_se,
               geom='errorbar',
               #fun.args=list(conf.int=.95),
               size=.6,
               aes(width=.3),
               color="black",
               position=pd) +
  coord_cartesian(ylim = c(1,7)) +
  scale_y_continuous(name = "Self-applicability", expand = c(0,0), breaks=seq(1,7,1)) + # set ticks at 1, 2, 3, 4 +
  scale_x_discrete(name = "Diversity Initiative Ad",
                   limits = c("majority black",
                              "majority white"),
                   labels = c("majority black" = "Majority\nBlack\nPanel",
                              "majority white" = "Majority\nWhite\nPanel")) +
  scale_fill_discrete(labels = c("black_p" = "Black women", 
                               "white_p" = "White women")) +
  theme_classic() +
  theme(text = element_text(color = "black"),
        axis.text.x = element_text(size = 14, color = "black"),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size = 14, color = "black"),
        axis.title.y = element_text(size = 14, color = "black"),
        #axis.title.y = element_blank(),
        legend.title = element_text(size = 18, hjust = .5),
        legend.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        legend.position = "top",
        legend.title.position = "top",
        #legend.justification = "center",
        #legend.direction = "vertical",
        ) + # +
  #labs(fill = "Participant race") +
  scale_colour_discrete(guide="none",
                        labels = c("black_p" = "Black women", "white_p" = "White women")
                        ) +
  scale_fill_manual(name = "Participant Race", 
                  labels = c("black_p" = "Black women", "white_p" = "White women"),
                  values = c("#2FB3EA", "#89939C")) +
  theme(plot.margin = margin(.5,.1,0,.25, "cm"))  #(order goes top, right, bottom, left) 

bsa
```

### Inclusion

```{r s3ainclusion}

pd = position_dodge(.92)

bi = data_bw %>%
  ggplot(aes(x = webad_race, y= in_mean, fill = prace_name)) +
  geom_bar(stat = "summary", 
           fun = mean,
           position = "dodge") +
  stat_summary(fun.data=mean_se,
               geom='errorbar',
               #fun.args=list(conf.int=.95),
               size=.6,
               aes(width=.3),
               color="black",
               position=pd) +
  coord_cartesian(ylim = c(1,7)) +
  scale_y_continuous(name = "Inclusion", expand = c(0,0), breaks=seq(1,7,1)) + # set ticks at 1, 2, 3, 4 +
  scale_x_discrete(name = "Diversity Initiative Ad",
                   limits = c("majority black",
                              "majority white"),
                   labels = c("majority black" = "Majority\nBlack\nPanel",
                              "majority white" = "Majority\nWhite\nPanel")) +
  scale_fill_discrete(labels = c("black_p" = "Black women", 
                               "white_p" = "White women")) +
  theme_classic() +
  theme(text = element_text(color = "black"),
        axis.text.x = element_text(size = 14, color = "black"),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size = 14, color = "black"),
        axis.title.y = element_text(size = 14, color = "black"),
        #axis.title.y = element_blank(),
        legend.title = element_text(size = 18, hjust = .5),
        legend.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        legend.position = "top",
        legend.title.position = "top",
        #legend.justification = "center",
        #legend.direction = "vertical",
        ) + # +
  #labs(fill = "Participant race") +
  scale_colour_discrete(guide="none",
                        labels = c("black_p" = "Black women", "white_p" = "White women")
                        ) +
  scale_fill_manual(name = "Participant Race", 
                  labels = c("black_p" = "Black women", "white_p" = "White women"),
                  values = c("#2FB3EA", "#89939C")) +
  theme(plot.margin = margin(.5,.1,0,.25, "cm"))  #(order goes top, right, bottom, left) 

bi
```

### Linked Fate

```{r s3alinkefate}

pd = position_dodge(.92)

blf = data_bw %>%
  ggplot(aes(x = webad_race, y= lfsb_mean, fill = prace_name)) +
  geom_bar(stat = "summary", 
           fun = mean,
           position = "dodge") +
  stat_summary(fun.data=mean_se,
               geom='errorbar',
               #fun.args=list(conf.int=.95),
               size=.6,
               aes(width=.3),
               color="black",
               position=pd) +
  coord_cartesian(ylim = c(1,7)) +
  scale_y_continuous(name = "Linked fate", expand = c(0,0), breaks=seq(1,7,1)) + # set ticks at 1, 2, 3, 4 +
  scale_x_discrete(name = "Diversity Initiative Ad",
                   limits = c("majority black",
                              "majority white"),
                   labels = c("majority black" = "Majority\nBlack\nPanel",
                              "majority white" = "Majority\nWhite\nPanel")) +
 # scale_fill_discrete(labels = c("black_p" = "Black women", 
  #                             "white_p" = "White women")) +
  theme_classic() +
  theme(text = element_text(color = "black"),
        axis.text.x = element_text(size = 14, color = "black"),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size = 14, color = "black"),
        axis.title.y = element_text(size = 14, color = "black"),
        #axis.title.y = element_blank(),
        legend.title = element_text(size = 18, hjust = .5),
        legend.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        legend.position = "top",
        legend.title.position = "top",
        #legend.justification = "center",
        #legend.direction = "vertical",
        ) + # +
  #labs(fill = "Participant race") +
  scale_colour_discrete(guide="none",
                        labels = c("black_p" = "Black women", "white_p" = "White women")
                        ) +
  scale_fill_manual(name = "Participant Race", 
                  labels = c("black_p" = "Black women", "white_p" = "White women"),
                  values = c("#2FB3EA", "#89939C")) +
  theme(plot.margin = margin(.5,.1,0,.25, "cm"))  #(order goes top, right, bottom, left) 

blf
```

### Study 3a Combined Figure

```{r s3aall}
ggarrange(blf,bsa,bi,
          nrow = 1,
          common.legend = TRUE)
```

## Study 3b: Latina Women and White Women

```{r s3bdta}

data_lw <- read_csv("~/Dropbox/Research/current projects/diversity gatekeeping/studies + data/workshop study Oct:Nov 24/data + analyses/lw_ww/gen div workshop LW_WW Nov 24 CLEAN.csv")

```

### Self-Applicability

```{r s3bsa}

pd = position_dodge(.92)

lsa = data_lw %>%
  ggplot(aes(x = webad_race, y= sa_mean, fill = prace_name)) +
  geom_bar(stat = "summary", 
           fun = mean,
           position = "dodge") +
  scale_y_continuous(name = "Self-applicability", expand = c(0,0), breaks=seq(1,7,1)) + # set ticks at 1, 2, 3, 4 +
  stat_summary(fun.data=mean_se,
               geom='errorbar',
               #fun.args=list(conf.int=.95),
               size=.6,
               aes(width=.3),
               color="black",
               position=pd) +
  coord_cartesian(ylim = c(1,7)) +
  scale_x_discrete(name = "Diversity Initiative Ad",
               limits = c("majority latina",
                          "majority white"),
               labels = c("majority latina" = "Majority\nLatina\nPanel",
                          "majority white" = "Majority\nWhite\nPanel")) +
# scale_fill_discrete(labels = c("black_p" = "Black women", 
#                             "white_p" = "White women")) +
  theme_classic() +
  theme(text = element_text(color = "black"),
        axis.text.x = element_text(size = 14, color = "black"),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size = 14, color = "black"),
        axis.title.y = element_text(size = 14, color = "black"),
        #axis.title.y = element_blank(),
        legend.title = element_text(size = 18, hjust = .5),
        legend.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        legend.position = "top",
        legend.title.position = "top",
        #legend.justification = "center",
        #legend.direction = "vertical",
        ) + # +
  #labs(fill = "Participant race") +
  scale_colour_discrete(guide="none",
                        labels = c("latina_p" = "Latina women", "white_p" = "White women")
                        ) +
  scale_fill_manual(name = "Participant Race", 
                  labels = c("latina_p" = "Latina women", "white_p" = "White women"),
                  values = c("#0078AB", "#89939C")) +
  theme(plot.margin = margin(.5,.1,0,.25, "cm"))  #(order goes top, right, bottom, left) 

lsa

```

### Inclusion

```{r s3binclusion}

pd = position_dodge(.92)

li = data_lw %>%
  ggplot(aes(x = webad_race, y= in_mean, fill = prace_name)) +
  geom_bar(stat = "summary", 
           fun = mean,
           position = "dodge") +
  stat_summary(fun.data=mean_se,
               geom='errorbar',
               #fun.args=list(conf.int=.95),
               size=.6,
               aes(width=.3),
               color="black",
               position=pd) +
  coord_cartesian(ylim = c(1,7)) +
  scale_y_continuous(name = "Inclusion", expand = c(0,0), breaks=seq(1,7,1)) + # set ticks at 1, 2, 3, 4 +
  scale_x_discrete(name = "Diversity Initiative Ad",
                 limits = c("majority latina",
                            "majority white"),
                 labels = c("majority latina" = "Majority\nLatina\nPanel",
                            "majority white" = "Majority\nWhite\nPanel")) +
# scale_fill_discrete(labels = c("black_p" = "Black women", 
#                             "white_p" = "White women")) +
  theme_classic() +
  theme(text = element_text(color = "black"),
        axis.text.x = element_text(size = 14, color = "black"),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size = 14, color = "black"),
        axis.title.y = element_text(size = 14, color = "black"),
        #axis.title.y = element_blank(),
        legend.title = element_text(size = 18, hjust = .5),
        legend.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        legend.position = "top",
        legend.title.position = "top",
        #legend.justification = "center",
        #legend.direction = "vertical",
        ) + # +
  #labs(fill = "Participant race") +
  scale_colour_discrete(guide="none",
                        labels = c("latina_p" = "Latina women", "white_p" = "White women")
                        ) +
  scale_fill_manual(name = "Participant Race", 
                  labels = c("latina_p" = "Latina women", "white_p" = "White women"),
                  values = c("#0078AB", "#89939C")) +
  theme(plot.margin = margin(.5,.1,0,.25, "cm"))  #(order goes top, right, bottom, left) 

li

```

### Linked Fate

```{r s3blinkefate}

pd = position_dodge(.92)

llf =  data_lw %>%
  ggplot(aes(x = webad_race, y= lfsb_mean, fill = prace_name)) +
  geom_bar(stat = "summary", 
           fun = mean,
           position = "dodge") +
  stat_summary(fun.data=mean_se,
               geom='errorbar',
               #fun.args=list(conf.int=.95),
               size=.6,
               aes(width=.3),
               color="black",
               position=pd) +
  coord_cartesian(ylim = c(1,7)) +
  scale_y_continuous(name = "Linked fate", expand = c(0,0), breaks=seq(1,7,1)) + # set ticks at 1, 2, 3, 4 +
  scale_x_discrete(name = "Diversity Initiative Ad",
                 limits = c("majority latina",
                            "majority white"),
                 labels = c("majority latina" = "Majority\nLatina\nPanel",
                            "majority white" = "Majority\nWhite\nPanel")) +
 # scale_fill_discrete(labels = c("black_p" = "Black women", 
  #                             "white_p" = "White women")) +
  theme_classic() +
  theme(text = element_text(color = "black"),
        axis.text.x = element_text(size = 14, color = "black"),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size = 14, color = "black"),
        axis.title.y = element_text(size = 14, color = "black"),
        #axis.title.y = element_blank(),
        legend.title = element_text(size = 18, hjust = .5),
        legend.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        legend.position = "top",
        legend.title.position = "top",
        #legend.justification = "center",
        #legend.direction = "vertical",
        ) + # +
  #labs(fill = "Participant race") +
  scale_colour_discrete(guide="none",
                        labels = c("latina_p" = "Latina women", "white_p" = "White women")
                        ) +
  scale_fill_manual(name = "Participant Race", 
                  labels = c("latina_p" = "Latina women", "white_p" = "White women"),
                  values = c("#0078AB", "#89939C")) +
  theme(plot.margin = margin(.5,.1,0,.25, "cm"))  #(order goes top, right, bottom, left) 

llf
```

### Study 3b Combined Figure

```{r s3bplots}
ggarrange(llf,lsa,li,
          nrow = 1,
          common.legend = TRUE)
```


