---
title: "Linked Fate Study 3b: Technical Appendix"
author: "Casey McMahon"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    code_folding: hide
    keep_md: true
---

# Study 3b: Latina and White Women Workshop Experiment

This technical appendix documents the Study 3b analyses for the linked fate workshop experiment. The study tested whether participants responded differently to gender-equality workshop advertisements depending on the racial composition of the panelists. The main outcomes were linked fate, self-applicability, and inclusion.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)

library(boot)
library(car)
library(corrr)
library(dplyr)
library(emmeans)
library(furniture)
library(ggplot2)
library(kableExtra)
library(knitr)
library(psych)
library(readr)
library(Rmisc)
library(stringr)
library(tibble)
library(tidyr)

clean_kable <- function(df, caption = NULL, align = NULL) {
  kableExtra::kable(
    df,
    format = "html",
    caption = caption,
    align = align,
    escape = FALSE
  ) %>%
    kableExtra::kable_styling(
      full_width = FALSE,
      position = "center",
      bootstrap_options = c("striped", "hover", "condensed")
    )
}

format_p <- function(p) {
  dplyr::case_when(
    is.na(p) ~ "",
    p < .001 ~ "< .001",
    TRUE ~ sprintf("%.3f", p)
  )
}

format_num <- function(x, digits = 2) {
  dplyr::if_else(is.na(x), "", sprintf(paste0("%.", digits, "f"), x))
}

clean_ad_label <- function(x) {
  dplyr::recode(
    as.character(x),
    "majority white" = "Majority White women panel",
    "majority black" = "Majority Black women panel",
    "majority latina" = "Majority Latina women panel",
    .default = as.character(x)
  )
}

clean_group_label <- function(x) {
  dplyr::recode(
    as.character(x),
    "Black" = "Black women",
    "White" = "White women",
    "Latina" = "Latina women",
    "Black women" = "Black women",
    "White women" = "White women",
    "Latina women" = "Latina women",
    .default = as.character(x)
  )
}

make_reliability_row <- function(scale_name, items) {
  alpha_obj <- psych::alpha(
    data %>% dplyr::select(dplyr::all_of(items)),
    check.keys = TRUE
  )

  tibble::tibble(
    Scale = scale_name,
    `Cronbach's alpha` = sprintf("%.2f", alpha_obj$total$raw_alpha)
  )
}

format_p_text <- function(p) {
  dplyr::case_when(
    is.na(p) ~ "",
    p < .001 ~ "< .001",
    TRUE ~ sprintf("%.3f", p)
  )
}

make_anova_table <- function(aov_model) {

  aov_df <- as.data.frame(summary(aov_model)[[1]]) %>%
    tibble::rownames_to_column("Effect") %>%
    dplyr::rename(
      df = Df,
      `Sum Sq` = `Sum Sq`,
      `Mean Sq` = `Mean Sq`,
      F = `F value`,
      p_raw = `Pr(>F)`
    )

  # Find the residual/error row. Different ANOVA objects label this differently.
  error_ss <- aov_df %>%
    dplyr::filter(stringr::str_detect(stringr::str_to_lower(Effect), "residual|error")) %>%
    dplyr::pull(`Sum Sq`)

  # If there is no residual row, leave partial eta squared blank rather than erroring.
  if (length(error_ss) == 0) {
    error_ss <- NA_real_
  } else {
    error_ss <- error_ss[1]
  }

  aov_df %>%
    dplyr::filter(!stringr::str_detect(stringr::str_to_lower(Effect), "residual|error")) %>%
    dplyr::mutate(
      Effect = stringr::str_replace_all(Effect, ":", " × "),
      F = format_num(F, 2),
      p = format_p_text(p_raw),
    `η<sub>p</sub>²` = if (is.na(error_ss)) {
      ""
    } else {
      format_num(`Sum Sq` / (`Sum Sq` + error_ss), 2)
    }
    ) %>%
    dplyr::select(
      Effect,
      df,
      F,
      p,
      `η<sub>p</sub>²`
    )
}

make_descriptive_table <- function(data, outcome) {
  data %>%
    dplyr::group_by(prace_name, webad_race) %>%
    dplyr::summarise(
      N = sum(!is.na(.data[[outcome]])),
      M = mean(.data[[outcome]], na.rm = TRUE),
      SD = sd(.data[[outcome]], na.rm = TRUE),
      .groups = "drop"
    ) %>%
    dplyr::mutate(
      `Participant race` = clean_group_label(prace_name),
      `Panel composition` = clean_ad_label(webad_race),
      M = format_num(M, 2),
      SD = format_num(SD, 2)
    ) %>%
    dplyr::select(`Participant race`, `Panel composition`, N, M, SD)
}

make_posthoc_table <- function(model) {
  emm <- emmeans::emmeans(model, specs = ~ prace_name * webad_race)
  contrast_obj <- pairs(emm, adjust = "tukey")

  contrast_table <- as.data.frame(contrast_obj)
  d_table <- as.data.frame(emmeans::eff_size(
    contrast_obj,
    sigma = sigma(model),
    edf = df.residual(model),
    method = "identity"
  ))

  contrast_table %>%
    dplyr::mutate(
      d = d_table$effect.size,
      Contrast = contrast,
      Contrast = stringr::str_replace_all(Contrast, "prace_name", "Participant race"),
      Contrast = stringr::str_replace_all(Contrast, "webad_race", "Panel composition"),
      Contrast = stringr::str_replace_all(Contrast, "Black(?! women)", "Black women"),
      Contrast = stringr::str_replace_all(Contrast, "Latina(?! women)", "Latina women"),
      Contrast = stringr::str_replace_all(Contrast, "White(?! women)", "White women"),
      Contrast = stringr::str_replace_all(Contrast, "majority white", "majority White women panel"),
      Contrast = stringr::str_replace_all(Contrast, "majority black", "majority Black women panel"),
      Contrast = stringr::str_replace_all(Contrast, "majority latina", "majority Latina women panel"),
      b = format_num(estimate, 2),
      SE = format_num(SE, 2),
      df = format_num(df, 1),
      t = format_num(t.ratio, 2),
      p = format_p(p.value),
      d = format_num(d, 2)
    ) %>%
    dplyr::select(Contrast, b, SE, df, t, p, d)
}

run_simple_mediation <- function(data, x, m, y, boot = 10000, seed = 654321) {
  analysis_data <- data %>%
    dplyr::select(dplyr::all_of(c(x, m, y))) %>%
    stats::na.omit()

  f_m <- stats::as.formula(paste(m, "~", x))
  f_y <- stats::as.formula(paste(y, "~", x, "+", m))
  f_total <- stats::as.formula(paste(y, "~", x))

  m_model <- stats::lm(f_m, data = analysis_data)
  y_model <- stats::lm(f_y, data = analysis_data)
  total_model <- stats::lm(f_total, data = analysis_data)

  a <- unname(stats::coef(m_model)[x])
  b <- unname(stats::coef(y_model)[m])
  c_prime <- unname(stats::coef(y_model)[x])
  c_total <- unname(stats::coef(total_model)[x])
  indirect <- a * b

  set.seed(seed)
  boot_fun <- function(d, indices) {
    bd <- d[indices, , drop = FALSE]
    bm <- stats::lm(f_m, data = bd)
    by <- stats::lm(f_y, data = bd)
    bt <- stats::lm(f_total, data = bd)
    ba <- unname(stats::coef(bm)[x])
    bb <- unname(stats::coef(by)[m])
    bc_prime <- unname(stats::coef(by)[x])
    bc_total <- unname(stats::coef(bt)[x])
    c(indirect = ba * bb, direct = bc_prime, total = bc_total)
  }

  boot_results <- boot::boot(analysis_data, statistic = boot_fun, R = boot)
  boot_mat <- boot_results$t
  colnames(boot_mat) <- c("indirect", "direct", "total")

  ci_indirect <- stats::quantile(boot_mat[, "indirect"], c(.025, .975), na.rm = TRUE)
  ci_direct <- stats::quantile(boot_mat[, "direct"], c(.025, .975), na.rm = TRUE)
  ci_total <- stats::quantile(boot_mat[, "total"], c(.025, .975), na.rm = TRUE)

  path_table <- tibble::tibble(
    Path = c("a path: participant race → linked fate", paste0("b path: linked fate → ", y), paste0("Indirect effect: participant race → linked fate → ", y)),
    Estimate = c(a, b, indirect),
    SE = c(summary(m_model)$coefficients[x, "Std. Error"], summary(y_model)$coefficients[m, "Std. Error"], stats::sd(boot_mat[, "indirect"], na.rm = TRUE)),
    `95% CI` = c("", "", paste0("[", sprintf("%.2f", ci_indirect[1]), ", ", sprintf("%.2f", ci_indirect[2]), "]"))
  ) %>%
    dplyr::mutate(
      Estimate = format_num(Estimate, 2),
      SE = format_num(SE, 2)
    )

  effects_table <- tibble::tibble(
    Effect = c("Total effect", "Direct effect", "Indirect effect"),
    Estimate = c(c_total, c_prime, indirect),
    SE = c(summary(total_model)$coefficients[x, "Std. Error"], summary(y_model)$coefficients[x, "Std. Error"], stats::sd(boot_mat[, "indirect"], na.rm = TRUE)),
    `95% CI` = c(
      paste0("[", sprintf("%.2f", ci_total[1]), ", ", sprintf("%.2f", ci_total[2]), "]"),
      paste0("[", sprintf("%.2f", ci_direct[1]), ", ", sprintf("%.2f", ci_direct[2]), "]"),
      paste0("[", sprintf("%.2f", ci_indirect[1]), ", ", sprintf("%.2f", ci_indirect[2]), "]")
    )
  ) %>%
    dplyr::mutate(
      Estimate = format_num(Estimate, 2),
      SE = format_num(SE, 2)
    )

  list(path_table = path_table, effects_table = effects_table)
}
```

```{r data-import, include = FALSE}
data <- readr::read_csv("~/Dropbox/Research/current projects/diversity gatekeeping/rmds + data for OSF/lf workshops/lw_ww/lf workshop_lw_ww_clean.csv")

if (!"ppid" %in% names(data)) {
  data <- data %>% tibble::rowid_to_column("ppid")
  data$ppid <- data$ppid + 100
}
```

## Overview

Study 3b tested whether Latina women and White women responded differently to a gender-equality workshop advertisement depending on whether the panel featured mostly Latina women or mostly White women. The primary outcomes were self-applicability, inclusion, and linked fate with the featured group.

## Sample and data checks

The visible sample table focuses on participant race because this study intentionally sampled women. I kept the gender check hidden in the source code as a data-cleaning check.

```{r hidden-gender-check, include=FALSE}
table(data$pgender, useNA = "ifany")
```

```{r sample-characteristics, echo=FALSE, results='asis'}
race_var <- "race_2"

sample_table <- data %>%
  dplyr::count(.data[[race_var]], name = "N") %>%
  dplyr::filter(!is.na(.data[[race_var]]), N > 0) %>%
  dplyr::mutate(
    Percent = sprintf("%.1f%%", 100 * N / sum(N)),
    `Participant group` = dplyr::case_when(
      .data[[race_var]] == 2 ~ "Latina women",
      .data[[race_var]] == 4 ~ "White women",
      TRUE ~ as.character(.data[[race_var]])
    )
  ) %>%
  dplyr::select(`Participant group`, N, Percent)

clean_kable(sample_table, caption = "Participant race")
```

```{r age-table, echo=FALSE, results='asis'}
age_data <- data %>%
  dplyr::mutate(age = as.numeric(age)) %>%
  dplyr::filter(!is.na(age), age > 10, age < 100)

age_table <- tibble::tibble(
  N = sum(!is.na(age_data$age)),
  M = mean(age_data$age, na.rm = TRUE),
  SD = sd(age_data$age, na.rm = TRUE),
  Min = min(age_data$age, na.rm = TRUE),
  Max = max(age_data$age, na.rm = TRUE)
) %>%
  dplyr::mutate(
    dplyr::across(M:Max, ~ sprintf("%.2f", .x))
  )

clean_kable(age_table, caption = "Participant age")
```

## Measures and scale construction

Participants responded to all items on 1–7 agreement scales. Higher scores indicate greater self-applicability, greater inclusion, and stronger linked fate.

<div class="measure-card">
  <h3>Self-applicability</h3>
  <ul>
    <li>I feel this workshop was intended for people like me.</li>
    <li>I believe that the workshop organizers were thinking of the experiences of people like me when they created this event.</li>
    <li>This workshop has the experiences of people like me in mind.</li>
  </ul>
</div>

<div class="measure-card">
  <h3>Inclusion</h3>
  <ul>
    <li>I would feel like I did not belong in this workshop. <em>Reverse-coded</em></li>
    <li>I would fit in the general environment of this workshop.</li>
    <li>I would fit in well with the other women at this workshop.</li>
    <li>I would feel out of place at this workshop. <em>Reverse-coded</em></li>
  </ul>
</div>

<div class="measure-card">
  <h3>Linked fate</h3>
  <ul>
    <li>In the workplace, I deal with similar issues as [target group].</li>
    <li>Workplace issues that affect me also affect [target group].</li>
    <li>Efforts to advance [target group] in the workplace will also help people like me to advance.</li>
    <li>Career progress for [target group] will also mean career progress for people like me.</li>
  </ul>
</div>

```{r scale-construction}
# Reverse-coded inclusion items
data <- data %>%
  dplyr::mutate(
    ex_1_r = dplyr::case_when(
      ex_1 == 1 ~ 7,
      ex_1 == 2 ~ 6,
      ex_1 == 3 ~ 5,
      ex_1 == 4 ~ 4,
      ex_1 == 5 ~ 3,
      ex_1 == 6 ~ 2,
      ex_1 == 7 ~ 1,
      TRUE ~ NA_real_
    ),
    ex_4_r = dplyr::case_when(
      ex_4 == 1 ~ 7,
      ex_4 == 2 ~ 6,
      ex_4 == 3 ~ 5,
      ex_4 == 4 ~ 4,
      ex_4 == 5 ~ 3,
      ex_4 == 6 ~ 2,
      ex_4 == 7 ~ 1,
      TRUE ~ NA_real_
    ),
    sa_mean = furniture::rowmeans(dplyr::select(., sa_1, sa_2, sa_3)),
    in_mean = furniture::rowmeans(dplyr::select(., ex_1_r, ex_2, ex_3, ex_4_r)),
    lf_mean = furniture::rowmeans(dplyr::select(., lfsb_1, lfsb_2, lfsb_3, lfsb_4))
  )

scale_items <- list(
  "Self-applicability" = c("sa_1", "sa_2", "sa_3"),
  "Inclusion" = c("ex_1_r", "ex_2", "ex_3", "ex_4_r"),
  "Linked fate" = c("lfsb_1", "lfsb_2", "lfsb_3", "lfsb_4")
)

scale_vars <- c("sa_mean", "in_mean", "lf_mean")
scale_labels <- c(
  "sa_mean" = "Self-applicability",
  "in_mean" = "Inclusion",
  "lf_mean" = "Linked fate"
)
```

The reliability table reports Cronbach's alpha for each scale. The correlations provide a quick check on how strongly the three outcomes are related.

```{r reliability-table, echo=FALSE, results='asis'}
reliability_table <- purrr::imap_dfr(
  scale_items,
  ~ make_reliability_row(.y, .x)
)

clean_kable(reliability_table, caption = "Scale reliability")
```

```{r scale-descriptives, echo=FALSE, results='asis'}
scale_descriptives <- data %>%
  dplyr::select(dplyr::all_of(scale_vars)) %>%
  tidyr::pivot_longer(dplyr::everything(), names_to = "Scale", values_to = "Score") %>%
  dplyr::group_by(Scale) %>%
  dplyr::summarise(
    N = sum(!is.na(Score)),
    M = mean(Score, na.rm = TRUE),
    SD = sd(Score, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::mutate(
    Scale = dplyr::recode(Scale, !!!scale_labels),
    M = format_num(M, 2),
    SD = format_num(SD, 2)
  )

clean_kable(scale_descriptives, caption = "Scale descriptive statistics")
```

```{r scale-correlations, echo=FALSE, results='asis'}
cor_mat <- data %>%
  dplyr::select(dplyr::all_of(scale_vars)) %>%
  stats::cor(use = "pairwise.complete.obs")

cor_mat[upper.tri(cor_mat, diag = TRUE)] <- NA

cor_table <- as.data.frame(round(cor_mat, 2)) %>%
  tibble::rownames_to_column("Scale") %>%
  dplyr::mutate(Scale = dplyr::recode(Scale, !!!scale_labels))

names(cor_table)[-1] <- dplyr::recode(names(cor_table)[-1], !!!scale_labels)

cor_table <- cor_table %>%
  dplyr::mutate(dplyr::across(-Scale, ~ dplyr::if_else(is.na(.x), "", sprintf("%.2f", .x))))

clean_kable(cor_table, caption = "Scale correlations")
```

## Primary ANOVAs

The primary analyses test whether each outcome differed by panel composition, participant race, and their interaction. The interaction term is the most important test because it indicates whether Latina women and White women responded differently to the majority-Latina versus majority-White workshop panels.

```{r prepare-analysis-vars}
data <- data %>%
  dplyr::mutate(
    prace_name = factor(prace_name),
    webad_race = factor(webad_race)
  )

sa_aov <- aov(sa_mean ~ webad_race * prace_name, data = data)
in_aov <- aov(in_mean ~ webad_race * prace_name, data = data)
lf_aov <- aov(lf_mean ~ webad_race * prace_name, data = data)
```

### Linked fate

This model tests whether participants felt more linked fate with panels that featured their own racial group than with panels featuring the other group.

```{r lf-anova-table, echo=FALSE, results='asis'}
clean_kable(make_anova_table(lf_aov), caption = "Linked fate ANOVA results")
```

```{r lf-descriptives, echo=FALSE, results='asis'}
clean_kable(make_descriptive_table(data, "lf_mean"), caption = "Linked fate means by participant race and panel composition")
```

```{r lf-posthoc, echo=FALSE, results='asis'}
clean_kable(make_posthoc_table(lf_aov), caption = "Linked fate Tukey-adjusted post hoc comparisons")
```

### Self-applicability

This model tests whether the workshop felt more applicable to participants when the panel featured mostly women from their own racial group.

```{r self-applicability-anova-table, echo=FALSE, results='asis'}
clean_kable(make_anova_table(sa_aov), caption = "Self-applicability ANOVA results")
```

```{r self-applicability-descriptives, echo=FALSE, results='asis'}
clean_kable(make_descriptive_table(data, "sa_mean"), caption = "Self-applicability means by participant race and panel composition")
```

```{r self-applicability-posthoc, echo=FALSE, results='asis'}
clean_kable(make_posthoc_table(sa_aov), caption = "Self-applicability Tukey-adjusted post hoc comparisons")
```

### Inclusion

This model tests whether the workshop environment felt more inclusive when the panel featured mostly women from the participant's own racial group.

```{r inclusion-anova-table, echo=FALSE, results='asis'}
clean_kable(make_anova_table(in_aov), caption = "Inclusion ANOVA results")
```

```{r inclusion-descriptives, echo=FALSE, results='asis'}
clean_kable(make_descriptive_table(data, "in_mean"), caption = "Inclusion means by participant race and panel composition")
```

```{r inclusion-posthoc, echo=FALSE, results='asis'}
clean_kable(make_posthoc_table(in_aov), caption = "Inclusion Tukey-adjusted post hoc comparisons")
```

Together, these analyses test whether representation in the workshop advertisement shaped linked fate, self-applicability, and inclusion differently for Latina women and White women.

## Mediation models

The mediation model tests whether linked fate explains why cross-race workshop panels were evaluated differently by Latina women and White women. The analysis uses only the cross-race workshop conditions: White women viewing the majority-Latina panel and Latina women viewing the majority-White panel. Participant race is coded 0 = White women and 1 = Latina women, so negative indirect effects indicate pathways that help explain lower self-applicability or inclusion among Latina women relative to White women in the cross-race conditions.

```{r mediation-code}
data <- data %>%
  dplyr::mutate(
    prace_process = dplyr::case_when(
      race_2 == 2 ~ 1,
      race_2 == 4 ~ 0,
      TRUE ~ NA_real_
    ),
    targ_gp_process = dplyr::case_when(
      webad_race == "majority latina" ~ 1,
      webad_race == "majority white" ~ 0,
      TRUE ~ NA_real_
    ),
    comparison = dplyr::case_when(
      prace_process == 0 & targ_gp_process == 0 ~ "ww_ww",
      prace_process == 0 & targ_gp_process == 1 ~ "ww_focal",
      prace_process == 1 & targ_gp_process == 0 ~ "focal_ww",
      prace_process == 1 & targ_gp_process == 1 ~ "focal_focal",
      TRUE ~ NA_character_
    )
  )

data_sub <- data %>%
  dplyr::filter(comparison %in% c("ww_focal", "focal_ww"))

sa_mediation_results <- run_simple_mediation(
  data = data_sub,
  x = "prace_process",
  m = "lf_mean",
  y = "sa_mean",
  boot = 10000,
  seed = 654321
)

in_mediation_results <- run_simple_mediation(
  data = data_sub,
  x = "prace_process",
  m = "lf_mean",
  y = "in_mean",
  boot = 10000,
  seed = 654321
)
```

### Self-applicability mediation

```{r self-applicability-mediation-tables, echo=FALSE, results='asis'}
clean_kable(sa_mediation_results$path_table, caption = "Self-applicability mediation paths and indirect effect")
clean_kable(sa_mediation_results$effects_table, caption = "Self-applicability total, direct, and indirect effects")
```

### Inclusion mediation

```{r inclusion-mediation-tables, echo=FALSE, results='asis'}
clean_kable(in_mediation_results$path_table, caption = "Inclusion mediation paths and indirect effect")
clean_kable(in_mediation_results$effects_table, caption = "Inclusion total, direct, and indirect effects")
```

The mediation tables show whether linked fate statistically accounts for differences in self-applicability and inclusion in the cross-race workshop conditions. Significant indirect effects indicate that lower linked fate helps explain why the same gender-equality workshop may feel less applicable or inclusive when the panel primarily features the other racial group.
