---
title: "DEI Commitment Update Study: Technical Appendix"
author: "Casey McMahon"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)

library(tidyverse)
library(psych)
library(broom)
library(emmeans)
library(kableExtra)
library(forcats)
library(patchwork)
```

# Overview

This appendix reports the main analyses in finer detail for the DEI commitment update study. Participants first read an initial DEI commitment and then read an update in which the organization either rolled back the commitment, maintained the original commitment, or doubled down on the commitment. The main question was whether organizational backtracking increased perceived hypocrisy and reduced perceptions of organizational credibility.

```{r load-data}
data <- readr::read_csv("data/dei-rollback-update-cleaned.csv")

# Use the binary-gender dataset only if needed for sensitivity checks.
# data <- readr::read_rds("data/dei-rollback-update-cleaned-binary-gender.rds")
```

# Helper functions

```{r helpers}
format_num <- function(x, digits = 2) {
  dplyr::if_else(
    is.na(x),
    "",
    sprintf(paste0("%.", digits, "f"), x)
  )
}

format_p_text <- function(p) {
  dplyr::case_when(
    is.na(p) ~ "",
    p < .001 ~ "< .001",
    TRUE ~ sprintf("%.3f", p)
  )
}

clean_kable <- function(df, caption = NULL, align = NULL, escape = FALSE) {
  kableExtra::kable(
    df,
    format = "html",
    caption = caption,
    align = align,
    escape = escape
  ) %>%
    kableExtra::kable_styling(
      full_width = FALSE,
      position = "center",
      bootstrap_options = c("striped", "hover", "condensed")
    )
}

#changing the reference group
data <- data %>%
  dplyr::mutate(
    condition_label = factor(
      condition_label,
      levels = c(
        "Rollback",
        "Maintained commitment",
        "Doubled down"
      )
    )
  )

clean_lm_table <- function(model, caption = NULL) {
  broom::tidy(model) %>%
    dplyr::mutate(
      Predictor = dplyr::case_when(
        term == "(Intercept)" ~ "Intercept",
        term == "condition_labelMaintained commitment" ~ "Condition: maintained commitment",
        term == "condition_labelDoubled down" ~ "Condition: doubled down",
        term == "race_simplePeople of color" ~ "Participant race: people of color",
        term == "gender_labelWomen" ~ "Participant gender: women",
        term == "gender_labelMen" ~ "Participant gender: men",
        term == "age" ~ "Age",
        TRUE ~ term
      ),
      b = sprintf("%.2f", estimate),
      SE = sprintf("%.2f", std.error),
      t = sprintf("%.2f", statistic),
      p = format_p_text(p.value)
    ) %>%
    dplyr::select(Predictor, b, SE, t, p) %>%
    clean_kable(caption = caption, align = c("l", "c", "c", "c", "c"), escape = TRUE)
}

make_alpha <- function(df, vars) {
  psych::alpha(df %>% dplyr::select(dplyr::all_of(vars)), check.keys = TRUE)$total$raw_alpha
}

make_descriptive_table <- function(df, outcome_vars, labels, group_var = condition_label) {
  df %>%
    dplyr::select({{ group_var }}, dplyr::all_of(outcome_vars)) %>%
    tidyr::pivot_longer(
      cols = dplyr::all_of(outcome_vars),
      names_to = "outcome",
      values_to = "value"
    ) %>%
    dplyr::group_by({{ group_var }}, outcome) %>%
    dplyr::summarise(
      N = sum(!is.na(value)),
      M = mean(value, na.rm = TRUE),
      SD = sd(value, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    dplyr::mutate(
      outcome = dplyr::recode(outcome, !!!labels),
      M = sprintf("%.2f", M),
      SD = sprintf("%.2f", SD)
    ) %>%
    dplyr::rename(
      Condition = {{ group_var }},
      Outcome = outcome
    )
}

format_p_html <- function(p) {
  dplyr::case_when(
    is.na(p) ~ "",
    p < .001 ~ "&lt; .001",
    TRUE ~ sprintf("%.3f", p)
  )
}

make_emmeans_table <- function(model, caption = NULL) {
  emm <- emmeans::emmeans(model, ~ condition_label)

  pairwise <- emmeans::contrast(
    emm,
    method = "pairwise"
  ) %>%
    summary(adjust = "tukey") %>%
    as.data.frame() %>%
    dplyr::mutate(
      p_clean = format_p_html(p.value),
      contrast = stringr::str_replace_all(contrast, " - ", " − ")
    ) %>%
    dplyr::select(
      Contrast = contrast,
      b = estimate,
      SE,
      df,
      t = t.ratio,
      p = p_clean
    ) %>%
    dplyr::mutate(
      b = sprintf("%.2f", b),
      SE = sprintf("%.2f", SE),
      df = sprintf("%.1f", df),
      t = sprintf("%.2f", t)
    )

  clean_kable(pairwise, caption = caption)
}

```

# Sample and data checks

The tables below summarize the analytic sample after exclusions and attention checks.

```{r sample-tables, echo=FALSE, results='asis'}
sample_condition <- data %>%
  dplyr::count(condition_label, name = "N") %>%
  dplyr::mutate(Percent = sprintf("%.1f%%", 100 * N / sum(N))) %>%
  dplyr::rename(Condition = condition_label)

clean_kable(sample_condition, caption = "Participants by condition", escape = TRUE)

sample_race <- data %>%
  dplyr::count(race_label, name = "N") %>%
  dplyr::filter(!is.na(race_label)) %>%
  dplyr::mutate(Percent = sprintf("%.1f%%", 100 * N / sum(N))) %>%
  dplyr::rename(`Race/ethnicity` = race_label)

clean_kable(sample_race, caption = "Participant race/ethnicity", escape = TRUE)

sample_gender <- data %>%
  dplyr::count(gender_label, name = "N") %>%
  dplyr::filter(!is.na(gender_label)) %>%
  dplyr::mutate(Percent = sprintf("%.1f%%", 100 * N / sum(N))) %>%
  dplyr::rename(Gender = gender_label)

clean_kable(sample_gender, caption = "Participant gender", escape = TRUE)

age_table <- data %>%
  dplyr::summarise(
    N = sum(!is.na(age)),
    M = mean(age, na.rm = TRUE),
    SD = sd(age, na.rm = TRUE),
    Min = min(age, na.rm = TRUE),
    Max = max(age, na.rm = TRUE)
  ) %>%
  dplyr::mutate(
    M = sprintf("%.2f", M),
    SD = sprintf("%.2f", SD)
  )

clean_kable(age_table, caption = "Participant age", escape = TRUE)
```

# Measures and scale reliability

The main outcomes were perceived hypocrisy, impressions of the company, perceived genuineness of the company’s diversity commitment, anticipated belonging, expectation violation, false signaling, perceived harm, and likelihood of recommending the organization to a friend from an underrepresented group.

```{r reliability-table, echo=FALSE, results='asis'}
reliability_table <- tibble::tibble(
  Scale = c(
    "Hypocrisy at Time 1",
    "Genuineness at Time 1",
    "Anticipated belonging at Time 1",
    "Hypocrisy at Time 2",
    "Anticipated belonging at Time 2",
    "Expectation violation",
    "False signaling"
  ),
  `Cronbach's alpha` = c(
    make_alpha(data, c("hyp_t1_1", "hyp_t1_2", "hyp_t1_3", "hyp_t1_4", "hyp_t1_5")),
    make_alpha(data, c("gen_t1_1", "gen_t1_2_r", "gen_t1_3_r", "gen_t1_4")),
    make_alpha(data, c("ab_t1_1", "ab_t1_2", "ab_t1_3", "ab_t1_4")),
    make_alpha(data, c("hyp_t2_1", "hyp_t2_2", "hyp_t2_3", "hyp_t2_4", "hyp_t2_5")),
    make_alpha(data, c("ab_t2_1", "ab_t2_2", "ab_t2_3", "ab_t2_4")),
    make_alpha(data, c("ev_1", "ev_2", "ev_3")),
    make_alpha(data, c("fs_d_1", "fs_d_2", "fs_u_3", "fs_u_4"))
  )
) %>%
  dplyr::mutate(`Cronbach's alpha` = sprintf("%.2f", `Cronbach's alpha`))

clean_kable(reliability_table, caption = "Scale reliability", escape = TRUE)
```

# Descriptive results by condition

Note: The key comparison is whether the rollback condition produced higher hypocrisy and lower credibility than maintaining or doubling down on the original DEI commitment.

```{r descriptive-tables, echo=FALSE, results='asis'}
outcome_labels <- c(
  hyp_t2_mean = "Hypocrisy",
  imp_t2 = "Company impression",
  ab_t2_mean = "Anticipated belonging",
  expectation_violation = "Expectation violation",
  false_signaling = "False signaling",
  recommend = "Recommendation likelihood"
)

descriptive_table_wide <- data %>%
  dplyr::select(condition_label, dplyr::all_of(names(outcome_labels))) %>%
  tidyr::pivot_longer(
    cols = dplyr::all_of(names(outcome_labels)),
    names_to = "outcome",
    values_to = "value"
  ) %>%
  dplyr::filter(!is.na(value), !is.na(condition_label)) %>%
  dplyr::group_by(outcome, condition_label) %>%
  dplyr::summarise(
    M = mean(value, na.rm = TRUE),
    SD = sd(value, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::mutate(
    outcome = dplyr::recode(outcome, !!!outcome_labels),
    stat = sprintf("%.2f (%.2f)", M, SD)
  ) %>%
  dplyr::select(outcome, condition_label, stat) %>%
  tidyr::pivot_wider(
    names_from = condition_label,
    values_from = stat
  ) %>%
  dplyr::rename(Outcome = outcome)

clean_kable(
  descriptive_table_wide,
  caption = "Descriptive statistics by condition, M (SD)",
  align = c("l", "c", "c", "c"),
  escape = TRUE
)
```

```{r condition-outcome-plot, echo=FALSE, fig.width=9, fig.height=5}
plot_df <- data %>%
  dplyr::select(condition_label, hyp_t2_mean, imp_t2, ab_t2_mean, expectation_violation, false_signaling, recommend) %>%
  tidyr::pivot_longer(
    cols = -condition_label,
    names_to = "outcome",
    values_to = "value"
  ) %>%
  dplyr::mutate(
    outcome = dplyr::recode(
      outcome,
      hyp_t2_mean = "Hypocrisy",
      imp_t2 = "Company impression",
      ab_t2_mean = "Anticipated belonging",
      expectation_violation = "Expectation violation",
      false_signaling = "False signaling",
      recommend = "Recommendation likelihood"
    )
  ) %>%
  dplyr::group_by(condition_label, outcome) %>%
  dplyr::summarise(
    N = sum(!is.na(value)),
    M = mean(value, na.rm = TRUE),
    SE = sd(value, na.rm = TRUE) / sqrt(N),
    ci_lower = M - qt(.975, df = N - 1) * SE,
    ci_upper = M + qt(.975, df = N - 1) * SE,
    .groups = "drop"
  )

rollback_update_outcomes_plot <- ggplot(
  plot_df,
  aes(x = condition_label, y = M, group = outcome, colour = condition_label)
) +
  geom_point(size = 2.2) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = 0.12, linewidth = 0.5) +
  facet_wrap(~ outcome, scales = "free_y") +
  labs(
    title = "Organizational rollback shaped perceptions of credibility",
    x = NULL,
    y = "Mean rating"
  ) +
  theme_classic(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold"),
    axis.text.x = element_text(angle = 25, hjust = 1)
  )

rollback_update_outcomes_plot
```

```{r export-condition-outcome-plot, eval=FALSE}
ggsave(
  filename = "../assets/img/dei-rollback-update-outcomes.png",
  plot = rollback_update_outcomes_plot,
  width = 9,
  height = 5,
  dpi = 300
)
```

# Main condition models

The models below test whether the commitment-update condition predicted the key Time 2 outcomes. (The rollback condition is the reference group, so positive or negative coefficients for the other conditions indicate how maintaining or doubling down differed from rollback.)

```{r main-models, echo=FALSE, results='asis'}
hyp_model <- lm(hyp_t2_mean ~ condition_label, data = data)
imp_model <- lm(imp_t2 ~ condition_label, data = data)
belonging_model <- lm(ab_t2_mean ~ condition_label, data = data)
ev_model <- lm(expectation_violation ~ condition_label, data = data)
fs_model <- lm(false_signaling ~ condition_label, data = data)
recommend_model <- lm(recommend ~ condition_label, data = data)

clean_lm_table(hyp_model, caption = "Condition model predicting hypocrisy")
clean_lm_table(imp_model, caption = "Condition model predicting company impressions")
clean_lm_table(belonging_model, caption = "Condition model predicting anticipated belonging")
clean_lm_table(ev_model, caption = "Condition model predicting expectation violation")
clean_lm_table(fs_model, caption = "Condition model predicting false signaling")
clean_lm_table(recommend_model, caption = "Condition model predicting recommendation likelihood")
```

# Post hoc condition comparisons

The post hoc tests compare the three commitment-update conditions directly to one another. 

```{r posthoc-tests, echo=FALSE, results='asis'}
make_emmeans_table(hyp_model, caption = "Post hoc comparisons for hypocrisy")
make_emmeans_table(imp_model, caption = "Post hoc comparisons for company impressions")
make_emmeans_table(belonging_model, caption = "Post hoc comparisons for anticipated belonging")
make_emmeans_table(ev_model, caption = "Post hoc comparisons for expectation violation")
make_emmeans_table(fs_model, caption = "Post hoc comparisons for false signaling")
make_emmeans_table(recommend_model, caption = "Post hoc comparisons for recommendation likelihood")
```

```{r export-figures, eval=FALSE, include = FALSE}
# Figures for portfolio
ggsave(
  filename = "~/assets/img/dei-rollback-update-plot.png",
  plot = rollback_update_outcomes_plot,
  width = 8,
  height = 5,
  dpi = 300
)

```
