---
title: "DEI Commitment Update Study: Private Cleaning File"
author: "Casey McMahon"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)

library(tidyverse)
library(janitor)
library(psych)
library(readr)
```

# Purpose

This private file imports, cleans, and prepares the raw Qualtrics files for the DEI commitment update study. It should not be uploaded publicly because it reads raw data and may contain private participant information.

The public technical appendix should begin from the cleaned `.rds` file created at the end of this script.

# Data import

Update these file paths to point to the raw Qualtrics exports on your computer.

```{r import-connect}
connect_file <- "~/Documents/reseach projects for websites/dei rollbacks/pilot 7b vignette sit FA25/data + analyses/dei hypocrisy 2 cell (imagine SIT) Oct 24_November 4, 2024_16.20.csv"

cnames_c <- readr::read_csv(connect_file, n_max = 0) %>%
  names()

data_raw_c <- readr::read_csv(
  connect_file,
  skip = 3,
  col_names = cnames_c
)

data_c <- data_raw_c %>%
  janitor::clean_names()
```

```{r import-prolific}
prolific_file <- "~/Documents/reseach projects for websites/dei rollbacks/pilot 7b vignette sit FA25/data + analyses/dei hypocrisy 2 cell (imagine SIT) Nov 24_prolific_November 11, 2024_13.33.csv"

cnames_p <- readr::read_csv(prolific_file, n_max = 0) %>%
  names()

data_raw_p <- readr::read_csv(
  prolific_file,
  skip = 3,
  col_names = cnames_p
)

data_p <- data_raw_p %>%
  janitor::clean_names()
```

# Harmonize columns and combine files

```{r harmonize-combine}
# Rename overlapping columns after clean_names().
# These were named morality_1, morality_2, and morality_5 in the export.
data_c <- data_c %>%
  dplyr::rename(
    ev_1 = morality_1,
    ev_2 = morality_2,
    ev_3 = morality_5
  )

data_p <- data_p %>%
  dplyr::rename(
    ev_1 = morality_1,
    ev_2 = morality_2,
    ev_3 = morality_5
  )

# Remove basic Qualtrics metadata columns if they exist.
cols_to_drop <- c(
  "start_date", "end_date", "status", "progress", "duration_in_seconds",
  "recorded_date", "user_language"
)

data_c <- data_c %>%
  dplyr::select(-dplyr::any_of(cols_to_drop))

data_p <- data_p %>%
  dplyr::select(-dplyr::any_of(cols_to_drop))

# Harmonize participant ID fields.
data_c <- data_c %>%
  dplyr::mutate(similar_study = -99) %>%
  dplyr::rename(plat_ppid = participant_id)

data_p <- data_p %>%
  dplyr::rename(plat_ppid = prolific_pid)

# Add missing columns to each dataset before binding.
missing_in_c <- setdiff(names(data_p), names(data_c))
missing_in_p <- setdiff(names(data_c), names(data_p))

data_c[missing_in_c] <- NA
data_p[missing_in_p] <- NA

data <- dplyr::bind_rows(
  data_c %>% dplyr::select(names(data_p)),
  data_p
)
```

# Exclusions and attention checks

```{r exclusions}
# Remove previews/test rows and participants who did not consent.
data <- data %>%
  dplyr::filter(distribution_channel == "anonymous") %>%
  dplyr::filter(consent == 1)

# Keep only valid experimental conditions.
data <- data %>%
  dplyr::filter(condition %in% c(-1, 0, 1))

# Full-stop attention check.
data <- data %>%
  dplyr::mutate(
    punctuation = stringr::str_to_lower(punctuation),
    punctuation_clean = dplyr::recode(
      punctuation,
      "a period" = "period",
      "a period." = "period",
      "peroid" = "period",
      "period" = "period",
      "period sign" = "period",
      "period." = "period",
      "this is a period." = "period",
      .default = NA_character_,
      .missing = NA_character_
    )
  ) %>%
  dplyr::filter(punctuation_clean == "period")

# Create participant ID for analysis.
data <- data %>%
  tibble::rowid_to_column("ppid") %>%
  dplyr::mutate(ppid = ppid + 100)
```

# Demographic recoding

```{r demographics}
# Race recoding. Participants who selected multiple race/ethnicity categories are coded as bi/multiracial.
data <- data %>%
  dplyr::mutate(
    race_raw = as.character(race),
    race_r = dplyr::recode(
      race_raw,
      "1" = 1,  # White
      "2" = 2,  # Asian
      "3" = 3,  # Black
      "4" = 4,  # Hispanic
      "5" = 5,  # Native American
      "6" = 6,  # Middle Eastern
      "7" = 7,  # bi/multiracial
      "8" = 8,  # another race
      "12" = 7, "13" = 7, "14" = 7, "15" = 7,
      "34" = 7, "35" = 7, "37" = 7, "67" = 7,
      "126" = 7, "136" = 7, "146" = 7,
      .default = NA_real_
    ),
    race_label = dplyr::case_when(
      race_r == 1 ~ "White",
      race_r == 2 ~ "Asian",
      race_r == 3 ~ "Black",
      race_r == 4 ~ "Hispanic",
      race_r == 5 ~ "Native American",
      race_r == 6 ~ "Middle Eastern",
      race_r == 7 ~ "Bi/multiracial",
      race_r == 8 ~ "Another race",
      TRUE ~ NA_character_
    ),
    race_simple = dplyr::case_when(
      race_r == 1 ~ "White",
      !is.na(race_r) ~ "People of color",
      TRUE ~ NA_character_
    ),
    gender_label = dplyr::case_when(
      gender == 1 ~ "Men",
      gender == 2 ~ "Women",
      gender == 3 ~ "Non-binary",
      gender == 4 ~ "Another gender",
      TRUE ~ NA_character_
    ),
    condition_label = dplyr::case_when(
      condition == -1 ~ "Rollback",
      condition == 0 ~ "Maintained commitment",
      condition == 1 ~ "Doubled down",
      TRUE ~ NA_character_
    ),
    condition_label = factor(
      condition_label,
      levels = c("Rollback", "Maintained commitment", "Doubled down")
    ),
    age = as.numeric(age)
  )
```

# Scale construction

```{r scale-construction}
row_mean <- function(df, vars) {
  rowMeans(df[, vars], na.rm = TRUE)
}

reverse_1_7 <- function(x) {
  dplyr::case_when(
    x == 1 ~ 7,
    x == 2 ~ 6,
    x == 3 ~ 5,
    x == 4 ~ 4,
    x == 5 ~ 3,
    x == 6 ~ 2,
    x == 7 ~ 1,
    TRUE ~ NA_real_
  )
}

data <- data %>%
  dplyr::mutate(
    gen_t1_2_r = reverse_1_7(gen_t1_2),
    gen_t1_3_r = reverse_1_7(gen_t1_3)
  ) %>%
  dplyr::mutate(
    hyp_t1_mean = row_mean(., c("hyp_t1_1", "hyp_t1_2", "hyp_t1_3", "hyp_t1_4", "hyp_t1_5")),
    gen_t1_mean = row_mean(., c("gen_t1_1", "gen_t1_2_r", "gen_t1_3_r", "gen_t1_4")),
    ab_t1_mean = row_mean(., c("ab_t1_1", "ab_t1_2", "ab_t1_3", "ab_t1_4")),

    hyp_t2_mean = row_mean(., c("hyp_t2_1", "hyp_t2_2", "hyp_t2_3", "hyp_t2_4", "hyp_t2_5")),
    ab_t2_mean = row_mean(., c("ab_t2_1", "ab_t2_2", "ab_t2_3", "ab_t2_4")),

    harm_white = harm_1 - 4,
    harm_employees_of_color = harm_2 - 4,
    expectation_violation = row_mean(., c("ev_1", "ev_2", "ev_3")),
    false_signaling = row_mean(., c("fs_d_1", "fs_d_2", "fs_u_3", "fs_u_4"))
  )
```

# Analysis dataset

```{r analysis-dataset}
analysis_df <- data %>%
  dplyr::select(
    ppid,
    condition,
    condition_label,
    race_r,
    race_label,
    race_simple,
    gender,
    gender_label,
    age,
    hyp_t1_mean,
    imp_t1,
    gen_t1_mean,
    ab_t1_mean,
    hyp_t2_mean,
    imp_t2,
    ab_t2_mean,
    harm_white,
    harm_employees_of_color,
    expectation_violation,
    false_signaling,
    recommend,
    dplyr::any_of(c(
      "hyp_t1_1", "hyp_t1_2", "hyp_t1_3", "hyp_t1_4", "hyp_t1_5",
      "gen_t1_1", "gen_t1_2_r", "gen_t1_3_r", "gen_t1_4",
      "ab_t1_1", "ab_t1_2", "ab_t1_3", "ab_t1_4",
      "hyp_t2_1", "hyp_t2_2", "hyp_t2_3", "hyp_t2_4", "hyp_t2_5",
      "ab_t2_1", "ab_t2_2", "ab_t2_3", "ab_t2_4",
      "ev_1", "ev_2", "ev_3",
      "fs_d_1", "fs_d_2", "fs_u_3", "fs_u_4"
    ))
  )

# Optional: restrict to participants identifying as men or women if you want to match the original models.
analysis_df_binary_gender <- analysis_df %>%
  dplyr::filter(gender_label %in% c("Men", "Women"))
```

# Save cleaned data

```{r save-cleaned-data}
dir.create("data", showWarnings = FALSE, recursive = TRUE)
readr::write_csv(analysis_df, "data/dei-rollback-update-cleaned.csv")
readr::write_csv(analysis_df_binary_gender, "data/dei-rollback-update-cleaned-binary-gender.csv")
```
